---
title: "Host-Parasite Meta-analysis"
author: "Amanda Vicente, Beatriz Willink, Dave Civitello"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: united
    highlight: tango
    number_sections: false
    toc: true
    toc_float: true
    toc_depth: 4
---

Load packages
```{r load-packages, message=FALSE, results='hide'}
x <-
  c(
    "metafor", "kableExtra")

lapply(x, function(y) {
  # check if installed, if not install
  if (!y %in% installed.packages()[, "Package"])
    install.packages(y)
  
  # load package
  try(require(y, character.only = T), silent = T)
})
```


Read in data
```{r read-data}
dat <- read.csv("../data/MA_dat_21-07-2021.csv", header = T, sep = ",")

# subset data that have a "variation type". We'll parse based on variation type below
dat <- dat[-which(dat$Variation.Type == ""),]
dat <- subset(dat, select=-c(Data_source,Comment)) 
colnames(dat)
  
```

# Data cleaning
*There might be more of this coming...*

## Gradient categories 
```{r}
# Toxin is a type of pollution
dat$Gradient.category <- gsub(pattern = "Toxin", replacement = "Pollution", x = dat$Gradient.category)

# exclude ambiguous/equivocal categories
dat <- dat[which(dat$Gradient.category != "Ecological"),]

# Check gradient categories 
levels(as.factor(dat$Gradient.category))
```

## Median, min, max and n to mean and SD
```{r}
non_par <- dat[which(dat$Variation.Type == "Median_Min_Max"),]

# Estimate mean for control and treatment from median, maximum and minimum
non_par$Estimated_Mean_C <- (non_par$Lwr_C + 2 *non_par$Mean_C + non_par$Upr_C) / 4
non_par$Estimated_Mean_X <- (non_par$Lwr_X + 2 *non_par$Mean_X + non_par$Upr_X) / 4

# Estimate SD from range and n
# First, read in Xi_N table from Wang et al. (2014) BMC Medical Research Methodology
Xi_table <- read.csv("../scripts/Xi_N_Table.csv", header = T, sep = ",") 

# Estimate SD using the approximation for each N under 50
for (i in 1:length(non_par$ID)) {
  non_par$Estimated_SD_C[i] <-
    (non_par$Upr_C[i] - non_par$Lwr_C[i]) / Xi_table$Xi_N[which(Xi_table$N == non_par$N_C[i])]
  non_par$Estimated_SD_X[i] <-
    (non_par$Upr_X[i] - non_par$Lwr_X[i]) / Xi_table$Xi_N[which(Xi_table$N == non_par$N_C[i])]
}

# overwrite dat

dat$Mean_C[which(dat$Variation.Type == "Median_Min_Max")] <- non_par$Estimated_Mean_C 
dat$Mean_X[which(dat$Variation.Type == "Median_Min_Max")] <- non_par$Estimated_Mean_X

dat$Variation_C[which(dat$Variation.Type == "Median_Min_Max")] <- non_par$Estimated_SD_C
dat$Variation_X[which(dat$Variation.Type == "Median_Min_Max")] <- non_par$Estimated_SD_X

dat$Variation.Type[which(dat$Variation.Type == "Median_Min_Max")] <- "SD"

# Sanity check
levels(as.factor(dat$Variation.Type))
```

## Median, min, max, and IQ range to mean
```{r}
non_par <- dat[which(dat$Variation.Type == "Median_IQ"),]

# Estimate mean for control and treatment from median, maximum and minimum
non_par$Estimated_Mean_C <- (non_par$Lwr_C + 2 *non_par$Mean_C + non_par$Upr_C) / 4
non_par$Estimated_Mean_X <- (non_par$Lwr_X + 2 *non_par$Mean_X + non_par$Upr_X) / 4

# overwrite dat
dat$Mean_C[which(dat$Variation.Type == "Median_Min_Max")] <- non_par$Estimated_Mean_C 
dat$Mean_X[which(dat$Variation.Type == "Median_Min_Max")] <- non_par$Estimated_Mean_X

dat$Variation.Type[which(dat$Variation.Type == "Median_IQ")] <- "IQ"

# Sanity check
levels(as.factor(dat$Variation.Type))
```

## subset effects with workable "variation types"
```{r}
nrow(dat)

accepted_var <- c("SE", "SD", "CI", "Wald CI", "IQ", "OR", "t", "Z", "Z_reg")
dat <- dat[which(dat$Variation.Type %in% accepted_var),]

nrow(dat)

# store sample size in an object
N <- nrow(dat)
```

## Standardize survival/mortality
```{r standardize-survival}
# which studies report mortality? 
mort <- grep(pattern = "[M,m]ort", x = dat$Trait)

# if the study reports mortality replace success for failure
for (i in mort) {
  dat$Success_C[i] <- dat$N_C[i] - dat$Success_C[i]
  dat$Success_X[i] <- dat$N_X[i] - dat$Success_X[i]
  dat$Trait[i] <- "Survival (converted)"
   }
```
# Getting effect sizes

## Calculate OR and approximate variance from contingency tables
```{r OR}
for (i in 1:N) {
  # no correction need if none of the categories are 0
  if (dat$Variation.Type[i] == "OR" &
      dat$Success_C[i] > 0 &
      dat$Success_X[i] > 0 &
      (dat$N_C[i] - dat$Success_C[i]) > 0 &
      (dat$N_X[i] - dat$Success_X[i]) > 0) {
    dat$OR[i] <-
      (dat$Success_X[i] * (dat$N_C[i] - dat$Success_C[i])) / (dat$Success_C[i] * (dat$N_X[i] - dat$Success_X[i]))
    dat$Log.OR[i] <- log(dat$OR[i])
    # approximate variance
    dat$Log.OR.v[i] <-
      1 / dat$Success_X[i] + 1 / (dat$N_X[i] - dat$Success_X[i]) + 1 / dat$Success_C[i] + 1 / (dat$N_C[i] - dat$Success_C[i])
  }
  else {
    # if at least one category is 0, we apply Yate's correction to avoid dividing by 0
    if (dat$Variation.Type[i] == "OR" &
        (dat$Success_C[i] == 0 |
        dat$Success_X[i] == 0 |
        (dat$N_C[i] - dat$Success_C[i]) == 0 |
        (dat$N_X[i] - dat$Success_X[i]) == 0)) {
      dat$OR[i] <-
        ((dat$Success_X[i] + 0.5) * (dat$N_C[i] - dat$Success_C[i] + 0.5)) / ((dat$Success_C[i] + 0.5) * (dat$N_X[i] - dat$Success_X[i] + 0.5))
      dat$Log.OR[i] <- log(dat$OR[i])
      # approximate variance
      dat$Log.OR.v[i] <-
        1 / (dat$Success_X[i] + 0.5)  + 1 / (dat$N_X[i] - dat$Success_X[i] + 0.5) + 1 / (dat$Success_C[i] + 0.5) + 1 / (dat$N_C[i] - dat$Success_C[i] + 0.5)
    }
    else{
      # if there is no OR data, make these columns NA
      dat$OR[i] <- NA
      dat$Log.OR[i] <- NA
      dat$Log.OR.v[i] <- NA
    }
  }
}
```


## Calculate SD for comparisons with continuous normally distributed variables
```{r SD}
# create SD variables
dat$SD_C <- vector(length = N)
dat$SD_X <- vector(length = N)

for (i in 1:N) {
  # Calculate SD if SE is reported
  if (dat$Variation.Type[i] == "SE") {
    dat$SD_C[i] <- dat$Variation_C[i] * sqrt(dat$N_C[i])
    dat$SD_X[i] <- dat$Variation_X[i] * sqrt(dat$N_X[i])
  } else {
    # calculate SD if a 95% confidence interval for a normal distribution is reported
    # this also applies to the Wald confidence interval for proportions, as it is a normal approximation to the binomial
    if (dat$Variation.Type[i] == "CI" |
        dat$Variation.Type[i] == "Wald CI") {
      # calculate SE from lower and upper limits
      temp_C1 <- (dat$Upr_C[i] - dat$Mean_C[i]) / 1.96
      temp_C2 <- (dat$Mean_C[i] - dat$Lwr_C[i]) / 1.96
      
      temp_X1 <- (dat$Upr_X[i] - dat$Mean_X[i]) / 1.96
      temp_X2 <- (dat$Mean_X[i] - dat$Lwr_X[i]) / 1.96
      
      # average digitized/recorded values of SE (because we have two) and transform to SD
      dat$SD_C[i] <-
        mean(abs(c(temp_C1, temp_C2))) * sqrt(dat$N_C[i])
      dat$SD_X[i] <-
        mean(abs(c(temp_C1, temp_C2))) * sqrt(dat$N_X[i])
    } else {
      # approximate SD if IQ range is reported
      if (dat$Variation.Type[i] == "IQ") {
        dat$SD_C[i] <- (dat$Upr_C[i] - dat$Lwr_C[i]) / 1.35
        dat$SD_X[i] <- (dat$Upr_X[i] - dat$Lwr_X[i]) / 1.35
      }
      else {
        # if SD is reported leave as such
        if (dat$Variation.Type[i] == "SD") {
          dat$SD_C[i] <- dat$Variation_C[i]
          dat$SD_X[i] <- dat$Variation_X[i]
        } else {
          # if there is no appropriate data, make these columns NA
          dat$SD_C[i] <- NA
          dat$SD_X[i] <- NA
        }
      }
    }
  }
}
```


## Get effect sizes and variances
```{r calculate d}
# within groups standard deviation
dat$S_within <-
  sqrt(((dat$N_C - 1) * dat$SD_C ^ 2 + (dat$N_X - 1) * dat$SD_X ^ 2) / (dat$N_C + dat$N_X - 2))

# standardized effect size
for (i in 1:N) {
  # standardized mean difference
  # we can't include data without variance
  if (is.na(dat$S_within[i]) == FALSE & dat$S_within[i] > 0) {
    dat$d[i] <- (dat$Mean_X[i] - dat$Mean_C[i]) / dat$S_within[i]
    dat$var.d[i] <-
      (dat$N_C[i] + dat$N_X[i]) / (dat$N_C[i] * dat$N_X[i]) + (dat$d[i] ^ 2) /
      (2 * (dat$N_C[i] + dat$N_X[i]))
  }
  else{
    # if only F statistic is reported - no cases of this yet!
    if (dat$Variation.Type[i] == "F_X") {
      dat$d[i] <-
        sqrt(dat$F_X[i] * (dat$N_X[i] + dat$N_C[i]) / (dat$N_C[i] * dat$N_X[i]))
      dat$var.d[i] <-
        (dat$N_C[i] + dat$N_X[i]) / (dat$N_C[i] * dat$N_X[i]) + (dat$d[i] ^ 2) /
        (2 * (dat$N_C[i] + dat$N_X[i]))
      
    } else{
      # if Z and N are reported for a regression analysis
      if (dat$Variation.Type[i] == "Z_reg") {
        r <- dat$Z_X[i] / sqrt(dat$Z_N[i])
        dat$d[i] <- (2 * r) / (sqrt(1 - r ^ 2))
        Vr <- (1 - r ^ 2) ^ 2 / (dat$Z_N[i] - 1)
        dat$var.d[i] <- (4 * Vr) / (1 - r ^ 2) ^ 3
      }
      else{
        # if t and N are reported for a regression analysis
        if (dat$Variation.Type[i] == "t") {
          r <- sqrt((dat$Z_X[i] ^ 2) / (dat$Z_X[i] ^ 2 + dat$Z_df[i]))
          dat$d[i] <- (2 * r) / (sqrt(1 - r ^ 2))
          Vr <- (1 - r ^ 2) ^ 2 / (dat$R_N[i] - 1)
          dat$var.d[i] <- (4 * Vr) / (1 - r ^ 2) ^ 3
        }
        else{
          # If Z is reported from a comparison of two independent groups
          if (dat$Variation.Type[i] == "Z") {
            dat$d[i] <-
              sqrt(dat$Z_X[i] * sqrt(dat$N_C[i] + dat$N_X[i]) / (1 - sqrt(
                dat$Z_X[i] ^ 2 * (dat$N_C[i] + dat$N_X[i]) ^ -1
              )))
            dat$var.d[i] <-
              (dat$N_C[i] + dat$N_X[i]) / (dat$N_C[i] * dat$N_X[i]) + (dat$d[i] ^ 2) /
              (2 * (dat$N_C[i] + dat$N_X[i]))
          }
          else{
            # converting log OR
            if (is.na(dat$Log.OR.v[i]) == FALSE &
                dat$Log.OR.v[i] > 0) {
              dat$d[i] <- dat$Log.OR[i] * (sqrt(3) / pi)
              dat$var.d[i] <- dat$Log.OR.v[i] * 3 / pi ^ 2
            } else {
              # for now, leave other types of effects as NA
              dat$d[i] <- NA
              dat$var.d[i] <- NA
            }
          }
        }
      }
    }
  }
}
```


## Correct for sample size and get g
```{r calculate g}
# sample size correction factor
for (i in 1:N) {
  # if odds ratio or comparison between means
  if (is.na(dat$N_C[i]) == F) {
    dat$J[i] <- 1 - 3 / (4 * (dat$N_C[i] + dat$N_X[i] - 2) - 1)
  } else {
    # if correlation
    if (is.na(dat$R_N[i]) == F) {
      dat$J[i] <- 1 - 3 / (4 * (dat$R_N[i] - 2) - 1)
    }else {
      dat$J[i] <- NA
    }
  }
}
# corrected effect size
dat$g <- dat$J * dat$d

# and variance
dat$var.g <- (dat$J ^ 2) * dat$var.d
```

# Prepare data
Filter experiments with both epidemiological and infected demographic effects
```{r divide-datasets}
# For now, get rid of NAs
dat <- dat[(is.na(dat$d)==F),]

# How many experiments do we have?
Exps <- unique(dat$Experiment)

# A data frame of experiments that included Infected demographic and Uninfected demographic effects
dat_1 <- data.frame()

# A data frame of experiments that included Infected demographic and Epidemiological effects
dat_2 <- data.frame()

# populate data sets with corresponding experiments
for (i in Exps) {
  tmp <- dat[which(dat$Experiment == i), ]
  if ("Uninfected demographic" %in% tmp$Trait.category &
      "Infected demographic" %in% tmp$Trait.category) {
    dat_1 <- rbind(dat_1, tmp)
  }
  if ("Epidemiological" %in% tmp$Trait.category &
      "Infected demographic" %in% tmp$Trait.category) {
    dat_2 <- rbind(dat_2, tmp)
  }
}

dat_1 <- dat_1[which(dat_1$Trait.category == "Uninfected demographic" | dat_1$Trait.category == "Infected demographic"),]

dat_2 <- dat_2[which(dat_2$Trait.category == "Epidemiological" | dat_2$Trait.category == "Infected demographic"),]
```

## Variance-covariance of sampling errors matrix (for multiple comparisons)
This code produces a variance-covariance of sampling errors matrix, n x n, with n = number of effect sizes. It requires the dataset (dat), Hedge's g (g), and the variance of Hedge's g (var.g) to be defined above. It uses "Experiment", "Level_C", "Trait", "N_C", and "N_X" columns.


### Identify negative eigen values
First, we find out which studies have negative eigen values in their variance-covar matrix. This means covariance is larger than variance and it's an indicator of suspiciously low variance and potentially an incorrectly reported N or an error while digitizing data.

For data set 1
```{r check-var-covar_1}
studies_1 <- unique(dat_1$Study)
studies_to_check <- c()
negative_eigen <- c()

for (k in studies_1) {
  dat_study <- dat_1[which (dat_1$Study == k),]
  varcovmat = matrix(0,
                     nrow = dim(dat_study)[1],
                     ncol = dim(dat_study)[1])
  
  for (i in 1:dim(dat_study)[1]) {
    for (j in 1:dim(dat_study)[1]) {
      if (i == j) {
        varcovmat[i, j] = dat_study$var.g[i]
      } else{
        if (dat_study[i, "Experiment"] == dat_study[j, "Experiment"] &
            dat_study[i, "Level_C"] == dat_study[j, "Level_C"] &
            dat_study[i, "Trait"] == dat_study[j, "Trait"] &
            dat_study[i, "Trait.category"] == dat_study[j, "Trait.category"]) {
          varcovmat[i, j] = 1 / dat_study[i, "N_C"] + dat_study$g[i] * dat_study$g[j] /
            (dat_study[i, "N_C"] + dat_study[i, "N_X"] + dat_study[j, "N_X"])
        }
      }
    }
  }
  
  val <- eigen(varcovmat)
  for (m in 1:length(val$values)){
    if (val$values[m] < 0) {
    studies_to_check <- append(studies_to_check, k)
    negative_eigen <- append(negative_eigen, val$values[m])
    }
  }
}
    
Neg_eigen_1 <- data.frame(study = studies_to_check, eigen = negative_eigen)

Neg_eigen_1 %>%
  kbl() %>%
  kable_material(c("striped", "hover"), full_width = F)
```

For data set 2
```{r check-var-covar_2}
studies_2 <- unique(dat_2$Study)
studies_to_check <- c()
negative_eigen <- c()

for (k in studies_2) {
  dat_study <- dat_2[which (dat_2$Study == k),]
  varcovmat = matrix(0,
                     nrow = dim(dat_study)[1],
                     ncol = dim(dat_study)[1])
  
  for (i in 1:dim(dat_study)[1]) {
    for (j in 1:dim(dat_study)[1]) {
      if (i == j) {
        varcovmat[i, j] = dat_study$var.g[i]
      } else{
        if (dat_study[i, "Experiment"] == dat_study[j, "Experiment"] &
            dat_study[i, "Level_C"] == dat_study[j, "Level_C"] &
            dat_study[i, "Trait"] == dat_study[j, "Trait"] &
            dat_study[i, "Trait.category"] == dat_study[j, "Trait.category"]) {
          varcovmat[i, j] = 1 / dat_study[i, "N_C"] + dat_study$g[i] * dat_study$g[j] /
            (dat_study[i, "N_C"] + dat_study[i, "N_X"] + dat_study[j, "N_X"])
        }
      }
    }
  }
  
  val <- eigen(varcovmat)
  for (m in 1:length(val$values)){
    if (val$values[m] < 0) {
    studies_to_check <- append(studies_to_check, k)
    negative_eigen <- append(negative_eigen, val$values[m])
    }
  }
}
    
Neg_eigen_2 <- data.frame(study = studies_to_check, eigen = negative_eigen)

Neg_eigen_2 %>%
  kbl() %>%
  kable_material(c("striped", "hover"), full_width = F)
```

Based on these eigen values, we tentatively remove the studies/experiments that cause the most trouble. Some we know have issues with reporting N, other we need to go back to and double check
```{r var-covar-matrix}
# exclude studies with negative eigen values that make the matrix not positive definite
dat_1 <- dat_1[-grep("Ashraf", dat_1$Study),]
dat_1 <- dat_1[-grep("Pochini", dat_1$Study),]
dat_1 <- dat_1[-grep("Shostak et al. 2015. Journal of Parasitology", dat_1$Study),]

# this results in negative eigen values, probably because differences between resource levels are stark and data (fecundity) is far from normally distributed, the problem persist after log 10 transforming, for now, I'm removing the experiment
dat_1 <- dat_1[-grep(121, dat_1$Experiment),]

varcovmat_1 = matrix(0, nrow=dim(dat_1)[1], ncol=dim(dat_1)[1])

for(i in 1:dim(dat_1)[1]){
  for(j in 1:dim(dat_1)[1]){
    if(i == j){varcovmat_1[i,j] = dat_1$var.g[i]}else{
      if(dat_1[i, "Experiment"] == dat_1[j, "Experiment"] & dat_1[i, "Level_C"] == dat_1[j, "Level_C"] & dat_1[i, "Trait"] == dat_1[j, "Trait"] & dat_1[i, "Trait.category"] == dat_1[j, "Trait.category"]){
        varcovmat_1[i,j] = 1/dat_1[i,"N_C"] + dat_1$g[i]*dat_1$g[j]/(dat_1[i,"N_C"] + dat_1[i,"N_X"] + dat_1[j,"N_X"]) 
      }
    }
  }
}

# exclude studies with negative eigen values that make the matrix not positive definite
dat_2 <- dat_2[-grep("Ashraf", dat_2$Study),]
dat_2 <- dat_2[-grep("Waki&Yoshinaga", dat_2$Study),]
dat_2 <- dat_2[-grep("Souto et al. 2015. VetMicrobiol.", dat_2$Study),]
#dat_2 <- dat_2[-grep("Purcell", dat_2$Study),]
#dat_2 <- dat_2[-grep("Bradley", dat_2$Study),]
dat_2 <- dat_2[-grep("Doublet", dat_2$Study),]
#dat_2 <- dat_2[-grep("Naggar and Paxton. 2020. Microbial Biotechnology", dat_2$Study),] 

varcovmat_2 = matrix(0, nrow=dim(dat_2)[1], ncol=dim(dat_2)[1])

for(i in 1:dim(dat_2)[1]){
  for(j in 1:dim(dat_2)[1]){
    if(i == j){varcovmat_2[i,j] = dat_2$var.g[i]}else{
      if(dat_2[i, "Experiment"] == dat_2[j, "Experiment"] & dat_2[i, "Level_C"] == dat_2[j, "Level_C"] & dat_2[i, "Trait"] == dat_2[j, "Trait"] & dat_2[i, "Trait.category"] == dat_2[j, "Trait.category"]){
        varcovmat_2[i,j] = 1/dat_2[i,"N_C"] + dat_2$g[i]*dat_2$g[j]/(dat_2[i,"N_C"] + dat_2[i,"N_X"] + dat_2[j,"N_X"]) 
      }
    }
  }
}  


# For now remove rare host types to check interaction with host
dat_2_byHost <-dat_2[which(dat_2$Host.type == "Arthropod" | dat_2$Host.type == "Mollusc" | dat_2$Host.type == "Fish" | dat_2$Host.type == "Amphibian"), ]

varcovmat_2_byHost = matrix(0, nrow=dim(dat_2_byHost)[1], ncol=dim(dat_2_byHost)[1])

for(i in 1:dim(dat_2_byHost)[1]){
  for(j in 1:dim(dat_2_byHost)[1]){
    if(i == j){varcovmat_2_byHost[i,j] = dat_2_byHost$var.g[i]}else{
      if(dat_2_byHost[i, "Experiment"] == dat_2_byHost[j, "Experiment"] & dat_2_byHost[i, "Level_C"] == dat_2_byHost[j, "Level_C"] & dat_2_byHost[i, "Trait"] == dat_2_byHost[j, "Trait"] & dat_2_byHost[i, "Trait.category"] == dat_2_byHost[j, "Trait.category"]){
        varcovmat_2_byHost[i,j] = 1/dat_2_byHost[i,"N_C"] + dat_2_byHost$g[i]*dat_2_byHost$g[j]/(dat_2_byHost[i,"N_C"] + dat_2_byHost[i,"N_X"] + dat_2_byHost[j,"N_X"]) 
      }
    }
  }
}
```

# Descriptive statistics
```{r descriptive-stats, echo=FALSE}
d_stats <-
  data.frame(
    stat = c(
      "Number of effects",
      "Number of experiments",
      "Number of arthropod species",
      "Number of mollusc species",
      "Number of fish species",
      "Number of amphibian species",
      "Number of reptile species",
      "Number of bird species",
      "Number of mammal species",
      "Number of viral taxa",
      "Number of bacterial taxa",
      "Number of Multiple infection",
      "Number of fungal taxa",
      "Number of protozan taxa",
      "Number of helminth taxa",
      "Number of environment experiments",
      "Number of pollution experiments",
      "Number of resource experiments",
      "Number of resistance effects",
      "Number of prevalence effects",
      "Number of intensity effects",
      "Number of tolerance effects",
      "Number of survival effects",
      "Number of fecundity effects"
    ),
    Question_1 = c(
      length(dat_1$g),
      length(unique(dat_1$Experiment)),
      length(unique(dat_1$Host[which(dat_1$Host.type == "Arthropod")])),
      length(unique(dat_1$Host[which(dat_1$Host.type == "Mollusc")])),
      length(unique(dat_1$Host[which(dat_1$Host.type == "Fish")])),
      length(unique(dat_1$Host[which(dat_1$Host.type == "Amphibian")])),
      length(unique(dat_1$Host[which(dat_1$Host.type == "Reptile")])),
      length(unique(dat_1$Host[which(dat_1$Host.type == "Bird")])),
      length(unique(dat_1$Host[which(dat_1$Host.type == "Mammal")])),
      length(unique(dat_1$Parasite[which(dat_1$Parasite.type ==
                                           "Virus")])),
      length(unique(dat_1$Parasite[which(dat_1$Parasite.type ==
                                           "Bacteria")])),
      length(unique(dat_1$Parasite[which(dat_1$Parasite.type ==
                                           "Multiple")])),
      length(unique(dat_1$Parasite[which(dat_1$Parasite.type ==
                                           "Fungus")])),
      length(unique(dat_1$Parasite[which(dat_1$Parasite.type ==
                                           "Protozoan")])),
      length(unique(dat_1$Parasite[which(dat_1$Parasite.type ==
                                           "Helminth")])),
      length(dat_1$Experiment[which(dat_1$Gradient.category ==
                                      "Environment")]),
      length(dat_1$Experiment[which(dat_1$Gradient.category ==
                                      "Pollution")]),
      length(dat_1$Experiment[which(dat_1$Gradient.category ==
                                      "Resource")]),
      length(dat_1$ID[which(dat_1$Trait.category == "Epidemiological")]),
      length(dat_1$ID[which(dat_1$Trait.type == "Prevalence")]),
      length(dat_1$ID[which(dat_1$Trait.type == "Intensity")]),
      length(dat_1$ID[grep("demographic", dat_1$Trait.category)]),
      length(dat_1$ID[which(dat_1$Trait.type == "Survivorship")]),
      length(dat_1$ID[which(dat_1$Trait.type == "Fecundity")])
    ),
    Question_2 = c(
      length(dat_2$g),
      length(unique(dat_2$Experiment)),
      length(unique(dat_2$Host[which(dat_2$Host.type == "Arthropod")])),
      length(unique(dat_2$Host[which(dat_2$Host.type == "Mollusc")])),
      length(unique(dat_2$Host[which(dat_2$Host.type == "Fish")])),
      length(unique(dat_2$Host[which(dat_2$Host.type == "Amphibian")])),
      length(unique(dat_2$Host[which(dat_2$Host.type == "Reptile")])),
      length(unique(dat_2$Host[which(dat_2$Host.type == "Bird")])),
      length(unique(dat_2$Host[which(dat_2$Host.type == "Mammal")])),
      length(unique(dat_2$Parasite[which(dat_2$Parasite.type ==
                                           "Virus")])),
      length(unique(dat_2$Parasite[which(dat_2$Parasite.type ==
                                           "Bacteria")])),
      length(unique(dat_2$Parasite[which(dat_2$Parasite.type ==
                                           "Multiple")])),
      length(unique(dat_2$Parasite[which(dat_2$Parasite.type ==
                                           "Fungus")])),
      length(unique(dat_2$Parasite[which(dat_2$Parasite.type ==
                                           "Protozoan")])),
      length(unique(dat_2$Parasite[which(dat_2$Parasite.type ==
                                           "Helminth")])),
      length(dat_2$Experiment[which(dat_2$Gradient.category ==
                                      "Environment")]),
      length(dat_2$Experiment[which(dat_2$Gradient.category ==
                                      "Pollution")]),
      length(dat_2$Experiment[which(dat_2$Gradient.category ==
                                      "Resource")]),
      length(dat_2$ID[which(dat_2$Trait.category == "Epidemiological")]),
      length(dat_2$ID[which(dat_2$Trait.type == "Prevalence")]),
      length(dat_2$ID[which(dat_2$Trait.type == "Intensity")]),
      length(dat_2$ID[grep("demographic", dat_2$Trait.category)]),
      length(dat_2$ID[which(dat_2$Trait.type == "Survivorship")]),
      length(dat_2$ID[which(dat_2$Trait.type == "Fecundity")])
    )
  )

d_stats %>%
  kbl() %>%
  kable_material(c("striped", "hover"), full_width = F)
```


# Preliminary analysis

## Question 1 - Are fitness traits negatively affected by stressors
### Null model
```{r Q1-null}
Q1m1 <-
  rma.mv(
    abs(g) ~ 1,
    V = varcovmat_1,
    random = list( ~ 1 |
                     ID, ~ 1 | Experiment, ~ 1 | Parasite),
    data = dat_1,
    method = "REML"
  )

summary(Q1m1)
```


## Single fixed effect
```{r Q1-single-fixed}
Q1m2 <-
  rma.mv(
    abs(g) ~ Trait.category -1,
    V = varcovmat_1,
    random = list( ~ 1 |
                     ID, ~ 1 | Experiment, ~ 1 | Parasite),
    data = dat_1,
    method = "REML"
  )

summary(Q1m2)
```

### Trait category x Gradient category interaction
```{r Q1-Trait-x-Gradient}
Q1m3 <-
  rma.mv(
    abs(g) ~ Trait.category:Gradient.category -1,
    V = varcovmat_1,
    random = list( ~ 1 |
                     ID, ~ 1 | Experiment, ~ 1 | Parasite),
    data = dat_1,
    method = "REML"
  )

summary(Q1m3)
```

### Trait category x Host type interaction
```{r Q1-Trait-x-Host}
Q1m4 <-
  rma.mv(
    abs(g) ~ Trait.category:Host.type -1,
    V = varcovmat_1,
    random = list( ~ 1 |
                     ID, ~ 1 | Experiment, ~ 1 | Parasite),
    data = dat_1,
    method = "REML"
  )

summary(Q1m4)
```

## Question 2 - Do resistance and tolerance traits have different sensitivity to stressors
### Null model
```{r Q2-null}
Q2m1 <-
  rma.mv(
    abs(g) ~ 1,
    V = varcovmat_2,
    random = list( ~ 1 |
                     ID, ~ 1 | Experiment, ~ 1 | Parasite),
    data = dat_2,
    method = "REML"
  )

summary(Q2m1)

```

## Single fixed effect
```{r Q2-single-fixed}
Q2m2 <-
  rma.mv(
    abs(g) ~ Trait.category -1,
    V = varcovmat_2,
    random = list( ~ 1 |
                     ID, ~ 1 | Experiment, ~ 1 | Parasite),
    data = dat_2,
    method = "REML"
  )

summary(Q2m2)
```

### Trait category x Gradient category interaction
```{r Q2-Trait-x-Gradient}
Q2m3 <-
  rma.mv(
    ab(g) ~ Trait.category:Gradient.category -1,
    V = varcovmat_2,
    random = list( ~ 1 |
                     ID, ~ 1 | Experiment, ~ 1 | Parasite),
    data = dat_2,
    method = "REML"
  )

summary(Q2m3)
```


### Trait category x Host type interaction
```{r Q2-Trait-x-Host}
Q2m4 <-
  rma.mv(
    abs(g) ~ Trait.category:Host.type -1,
    V = varcovmat_2_byHost,
    random = list( ~ 1 |
                     ID, ~ 1 | Experiment, ~ 1 | Parasite),
    data = dat_2_byHost,
    method = "REML",
)

summary(Q2m4)
```