---
title: "Host-Parasite Meta-analysis"
author: "Amanda Vicente, Beatriz Willink, Kacy Nowak, Dave Civitello"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: united
    highlight: tango
    number_sections: false
    toc: true
    toc_float: true
    toc_depth: 4
---

Load packages
```{r load-packages, message=FALSE, results='hide'}
x <-
  c(
    "metafor", "kableExtra", "Matrix", "dplyr", "wesanderson", "weightr", "ggplot2", "MuMIn", "glmulti")

lapply(x, function(y) {
  # check if installed, if not install
  if (!y %in% installed.packages()[, "Package"])
    install.packages(y)
  
  # load package
  try(require(y, character.only = T), silent = T)
})
```
*Note* If you have trouble loading the glmulti package due to issues with Java, follow https://zhiyzuo.github.io/installation-rJava/ using the version Java17 (make sure to download the Arm 64 DMG Installer at https://www.oracle.com/java/technologies/downloads/#jdk17-mac)

Read in data
```{r read-data}
dat <- read.csv("../data/MA_dat_08-05-2021.csv", header = T, sep = ",")

# subset data that have a "variation type". We'll parse based on variation type below
#dat <- dat[-which(is.na(dat$Variation.Type) == TRUE),]
dat <- subset(dat, select = -c(Data_source,Comment)) 
colnames(dat)
  
```

# Data cleaning
*There might be more of this coming...*

## Gradient categories 
```{r}
# Toxin is a type of pollution
dat$Gradient.category <- gsub(pattern = "Toxin", replacement = "Pollution", x = dat$Gradient.category)

# exclude ambiguous/equivocal categories
dat <- dat[which(dat$Gradient.category != "Ecological"),]

# Check gradient categories 
levels(as.factor(dat$Gradient.category))
```

## Median, min, max and n to mean and SD
```{r}
non_par <- dat[which(dat$Variation.Type == "Median_Min_Max"),]

# Estimate mean for control and treatment from median, maximum and minimum
non_par$Estimated_Mean_C <- (non_par$Lwr_C + 2 * non_par$Mean_C + non_par$Upr_C) / 4
non_par$Estimated_Mean_X <- (non_par$Lwr_X + 2 * non_par$Mean_X + non_par$Upr_X) / 4

# Estimate SD from range and n
# First, read in Xi_N table from Wang et al. (2014) BMC Medical Research Methodology
Xi_table <- read.csv("../scripts/Xi_N_Table.csv", header = T, sep = ",") 

# Estimate SD using the approximation for each N under 50
for (i in 1:length(non_par$ID)) {
  non_par$Estimated_SD_C[i] <-
    (non_par$Upr_C[i] - non_par$Lwr_C[i]) / Xi_table$Xi_N[which(Xi_table$N == non_par$N_C[i])]
  non_par$Estimated_SD_X[i] <-
    (non_par$Upr_X[i] - non_par$Lwr_X[i]) / Xi_table$Xi_N[which(Xi_table$N == non_par$N_X[i])]
}

# overwrite dat
dat$Mean_C[which(dat$Variation.Type == "Median_Min_Max")] <- non_par$Estimated_Mean_C 
dat$Mean_X[which(dat$Variation.Type == "Median_Min_Max")] <- non_par$Estimated_Mean_X

dat$Variation_C[which(dat$Variation.Type == "Median_Min_Max")] <- non_par$Estimated_SD_C
dat$Variation_X[which(dat$Variation.Type == "Median_Min_Max")] <- non_par$Estimated_SD_X

dat$Variation.Type[which(dat$Variation.Type == "Median_Min_Max")] <- "SD"

# Sanity check
levels(as.factor(dat$Variation.Type))
```

## Median and IQ range to mean 
```{r}
non_par <- dat[which(dat$Variation.Type == "Median_IQ"),]

# Estimate mean for control and treatment from median and IQ range
non_par$Estimated_Mean_C <- (non_par$Lwr_C + non_par$Mean_C + non_par$Upr_C) / 3
non_par$Estimated_Mean_X <- (non_par$Lwr_X + non_par$Mean_X + non_par$Upr_X) / 3

# overwrite dat
dat$Mean_C[which(dat$Variation.Type == "Median_IQ")] <- non_par$Estimated_Mean_C 
dat$Mean_X[which(dat$Variation.Type == "Median_IQ")] <- non_par$Estimated_Mean_X

dat$Variation.Type[which(dat$Variation.Type == "Median_IQ")] <- "IQ"

# Sanity check
levels(as.factor(dat$Variation.Type))
```

## subset effects with workable "variation types"
```{r}
nrow(dat)

accepted_var <- c("SE", "SD", "CI", "Wald CI", "IQ", "OR")
dat <- dat[which(dat$Variation.Type %in% accepted_var),]

nrow(dat)

# store sample size in an object
N <- nrow(dat)
```

# Getting effect sizes

## Calculate OR and approximate variance from contingency tables
```{r OR}
for (i in 1:N) {
  # no correction need if none of the categories are 0
  if (dat$Variation.Type[i] == "OR" &
      dat$Success_C[i] > 0 &
      dat$Success_X[i] > 0 &
      (dat$N_C[i] - dat$Success_C[i]) > 0 &
      (dat$N_X[i] - dat$Success_X[i]) > 0) {
    dat$OR[i] <-
      (dat$Success_X[i] * (dat$N_C[i] - dat$Success_C[i])) / (dat$Success_C[i] * (dat$N_X[i] - dat$Success_X[i]))
    dat$Log.OR[i] <- log(dat$OR[i])
    # approximate variance
    dat$Log.OR.v[i] <-
      1 / dat$Success_X[i] + 1 / (dat$N_X[i] - dat$Success_X[i]) + 1 / dat$Success_C[i] + 1 / (dat$N_C[i] - dat$Success_C[i])
  }
  else {
    # if at least one category is 0, we apply Yate's correction to avoid dividing by 0
    if (dat$Variation.Type[i] == "OR" &
        (dat$Success_C[i] == 0 |
        dat$Success_X[i] == 0 |
        (dat$N_C[i] - dat$Success_C[i]) == 0 |
        (dat$N_X[i] - dat$Success_X[i]) == 0)) {
      dat$OR[i] <-
        ((dat$Success_X[i] + 0.5) * (dat$N_C[i] - dat$Success_C[i] + 0.5)) / ((dat$Success_C[i] + 0.5) * (dat$N_X[i] - dat$Success_X[i] + 0.5))
      dat$Log.OR[i] <- log(dat$OR[i])
      # approximate variance
      dat$Log.OR.v[i] <-
        1 / (dat$Success_X[i] + 0.5)  + 1 / (dat$N_X[i] - dat$Success_X[i] + 0.5) + 1 / (dat$Success_C[i] + 0.5) + 1 / (dat$N_C[i] - dat$Success_C[i] + 0.5)
    }
    else{
      # if there is no OR data, make these columns NA
      dat$OR[i] <- NA
      dat$Log.OR[i] <- NA
      dat$Log.OR.v[i] <- NA
    }
  }
}
```


## Calculate SD for comparisons with continuous normally distributed variables
```{r SD}
# create SD variables
dat$SD_C <- vector(length = N)
dat$SD_X <- vector(length = N)

for (i in 1:N) {
  # Calculate SD if SE is reported
  if (dat$Variation.Type[i] == "SE") {
    dat$SD_C[i] <- dat$Variation_C[i] * sqrt(dat$N_C[i])
    dat$SD_X[i] <- dat$Variation_X[i] * sqrt(dat$N_X[i])
  } else {
    # calculate SD if a 95% confidence interval for a normal distribution is reported
    # this also applies to the Wald confidence interval for proportions, as it is a normal approximation to the binomial
    if (dat$Variation.Type[i] == "CI" |
        dat$Variation.Type[i] == "Wald CI") {
      # calculate SE from lower and upper limits
      temp_C1 <- (dat$Upr_C[i] - dat$Mean_C[i]) / 1.96
      temp_C2 <- (dat$Mean_C[i] - dat$Lwr_C[i]) / 1.96
      
      temp_X1 <- (dat$Upr_X[i] - dat$Mean_X[i]) / 1.96
      temp_X2 <- (dat$Mean_X[i] - dat$Lwr_X[i]) / 1.96
      
      # average digitized/recorded values of SE (because we have two) and transform to SD
      dat$SD_C[i] <-
        mean(abs(c(temp_C1, temp_C2))) * sqrt(dat$N_C[i])
      dat$SD_X[i] <-
        mean(abs(c(temp_C1, temp_C2))) * sqrt(dat$N_X[i])
    } else {
      # approximate SD if IQ range is reported
      if (dat$Variation.Type[i] == "IQ") {
        dat$SD_C[i] <- (dat$Upr_C[i] - dat$Lwr_C[i]) / 1.35
        dat$SD_X[i] <- (dat$Upr_X[i] - dat$Lwr_X[i]) / 1.35
      }
      else {
        # if SD is reported leave as such
        if (dat$Variation.Type[i] == "SD") {
          dat$SD_C[i] <- dat$Variation_C[i]
          dat$SD_X[i] <- dat$Variation_X[i]
        } else {
          # if there is no appropriate data, make these columns NA
          dat$SD_C[i] <- NA
          dat$SD_X[i] <- NA
        }
      }
    }
  }
}
```


## Get effect sizes and variances
```{r calculate d}
# within groups standard deviation
dat$S_within <-
  sqrt(((dat$N_C - 1) * dat$SD_C ^ 2 + (dat$N_X - 1) * dat$SD_X ^ 2) / (dat$N_C + dat$N_X - 2))

# standardized effect size
for (i in 1:N) {
  # standardized mean difference
  # we can't include data without variance
  if (is.na(dat$S_within[i]) == FALSE & dat$S_within[i] > 0) {
    dat$d[i] <- (dat$Mean_X[i] - dat$Mean_C[i]) / dat$S_within[i]
    dat$var.d[i] <-
      (dat$N_C[i] + dat$N_X[i]) / (dat$N_C[i] * dat$N_X[i]) + (dat$d[i] ^ 2) /
      (2 * (dat$N_C[i] + dat$N_X[i]))
  }
  else{
    # if only F statistic is reported - no cases of this yet!
    if (dat$Variation.Type[i] == "F_X") {
      dat$d[i] <-
        sqrt(dat$F_X[i] * (dat$N_X[i] + dat$N_C[i]) / (dat$N_C[i] * dat$N_X[i]))
      dat$var.d[i] <-
        (dat$N_C[i] + dat$N_X[i]) / (dat$N_C[i] * dat$N_X[i]) + (dat$d[i] ^ 2) /
        (2 * (dat$N_C[i] + dat$N_X[i]))
      
    } else{
      # if Z and N are reported for a regression analysis
      if (dat$Variation.Type[i] == "Z_reg") {
        r <- dat$Z_X[i] / sqrt(dat$Z_N[i])
        dat$d[i] <- (2 * r) / (sqrt(1 - r ^ 2))
        Vr <- (1 - r ^ 2) ^ 2 / (dat$Z_N[i] - 1)
        dat$var.d[i] <- (4 * Vr) / (1 - r ^ 2) ^ 3
      }
      else{
        # if t and N are reported for a regression analysis
        if (dat$Variation.Type[i] == "t") {
          r <- sqrt((dat$Z_X[i] ^ 2) / (dat$Z_X[i] ^ 2 + dat$Z_df[i]))
          dat$d[i] <- (2 * r) / (sqrt(1 - r ^ 2))
          Vr <- (1 - r ^ 2) ^ 2 / (dat$R_N[i] - 1)
          dat$var.d[i] <- (4 * Vr) / (1 - r ^ 2) ^ 3
        }
        else{
          # If Z is reported from a comparison of two independent groups
          if (dat$Variation.Type[i] == "Z") {
            dat$d[i] <-
              sqrt(abs(dat$Z_X[i]) * sqrt(dat$N_C[i] + dat$N_X[i]) / (1 - sqrt(
                dat$Z_X[i] ^ 2 * (dat$N_C[i] + dat$N_X[i]) ^ -1
              )))
            dat$var.d[i] <-
              (dat$N_C[i] + dat$N_X[i]) / (dat$N_C[i] * dat$N_X[i]) + (dat$d[i] ^ 2) /
              (2 * (dat$N_C[i] + dat$N_X[i]))
          }
          else{
            # converting log OR
            if (is.na(dat$Log.OR.v[i]) == FALSE &
                dat$Log.OR.v[i] > 0) {
              dat$d[i] <- dat$Log.OR[i] * (sqrt(3) / pi)
              dat$var.d[i] <- dat$Log.OR.v[i] * 3 / pi ^ 2
            } else {
              # for now, leave other types of effects as NA
              dat$d[i] <- NA
              dat$var.d[i] <- NA
            }
          }
        }
      }
    }
  }
}
```


## Correct for sample size and get g
```{r calculate g}
# sample size correction factor
for (i in 1:N) {
  # if odds ratio or comparison between means
  if (is.na(dat$N_C[i]) == F) {
    dat$J[i] <- 1 - 3 / (4 * (dat$N_C[i] + dat$N_X[i] - 2) - 1)
  } else {
    # if correlation
    if (is.na(dat$R_N[i]) == F) {
      dat$J[i] <- 1 - 3 / (4 * (dat$R_N[i] - 2) - 1)
    }else {
      dat$J[i] <- NA
    }
  }
}
# corrected effect size
dat$g <- dat$J * dat$d

# and variance
dat$var.g <- (dat$J ^ 2) * dat$var.d
```

## flip effect sizes for remaining mortality effects
```{r standardize-survival}
mort <- grep(pattern = "[M,m]ort", x = dat$Trait)

for(i in mort){
  dat$g[i] <- -dat$g[i]
}
```

# Prepare data
Filter experiments with both epidemiological and infected demographic effects
```{r divide-datasets}
# For now, get rid of NAs
dat <- dat[(is.na(dat$d) == F),]

# How many experiments do we have?
Exps <- unique(dat$Experiment)

# A data frame of experiments that included Infected demographic and Uninfected demographic effects
dat_1 <- data.frame()

# A data frame of experiments that included Infected demographic and Epidemiological effects
dat_3 <- data.frame()

# populate data sets with corresponding experiments
for (i in Exps) {
  tmp <- dat[which(dat$Experiment == i), ]
  if ("Uninfected demographic" %in% tmp$Trait.category &
      "Infected demographic" %in% tmp$Trait.category) {
    dat_1 <- rbind(dat_1, tmp)
  }
  if ("Epidemiological" %in% tmp$Trait.category &
      "Infected demographic" %in% tmp$Trait.category) {
    dat_3 <- rbind(dat_3, tmp)
  }
}

dat_1 <- dat_1[which(dat_1$Trait.category == "Uninfected demographic" | dat_1$Trait.category == "Infected demographic"),]

dat_3 <- dat_3[which(dat_3$Trait.category == "Epidemiological" | dat_3$Trait.category == "Infected demographic"),]

# A data frame with experiments that include just epidemiological effects
dat_2 <- dat_3[which(dat_3$Trait.category == "Epidemiological"),]
```

## Variance-covariance of sampling errors matrix (for multiple comparisons)
This code produces a variance-covariance of sampling errors matrix, n x n, with n = number of effect sizes. It requires the dataset (dat), Hedge's g (g), and the variance of Hedge's g (var.g) to be defined above. It uses "Experiment", "Level_C", "Trait", "N_C", and "N_X" columns.


### Identify negative eigenvalues
First, we find out which studies have negative eigenvalues in their variance-covar matrix. This means covariance between treatment effects with the same control is larger than variance and it's an indicator of suspiciously low variance and potentially an incorrectly reported N or an error while digitizing data.

For data set 1
```{r check-var-covar_1}
studies_1 <- unique(dat_1$Study)
studies_to_check <- c()
negative_eigen <- c()

for (k in studies_1) {
  dat_study <- dat_1[which(dat_1$Study == k),]
  varcovmat = matrix(0,
                     nrow = dim(dat_study)[1],
                     ncol = dim(dat_study)[1])
  
  for (i in 1:dim(dat_study)[1]) {
    for (j in 1:dim(dat_study)[1]) {
      if (i == j) {
        varcovmat[i, j] = dat_study$var.g[i]
      } else{
        if (dat_study[i, "Experiment"] == dat_study[j, "Experiment"] &
            dat_study[i, "Level_C"] == dat_study[j, "Level_C"] &
            dat_study[i, "Trait"] == dat_study[j, "Trait"] &
            dat_study[i, "Trait.category"] == dat_study[j, "Trait.category"]) {
          varcovmat[i, j] = 1 / dat_study[i, "N_C"] + dat_study$g[i] * dat_study$g[j] /
            (dat_study[i, "N_C"] + dat_study[i, "N_X"] + dat_study[j, "N_X"])
        }
      }
    }
  }
  
  val <- eigen(varcovmat)
  for (m in 1:length(val$values)) {
    if (val$values[m] < 0) {
    studies_to_check <- append(studies_to_check, k)
    negative_eigen <- append(negative_eigen, val$values[m])
    }
  }
}
    
Neg_eigen_1 <- data.frame(study = studies_to_check, eigen = negative_eigen)

Neg_eigen_1 %>%
  kbl() %>%
  kable_material(c("striped", "hover"), full_width = F)
```
For data set 2
```{r check-var-covar_2}
studies_2 <- unique(dat_2$Study)
studies_to_check <- c()
negative_eigen <- c()

for (k in studies_2) {
  dat_study <- dat_2[which(dat_2$Study == k),]
  varcovmat = matrix(0,
                     nrow = dim(dat_study)[1],
                     ncol = dim(dat_study)[1])
  
  for (i in 1:dim(dat_study)[1]) {
    for (j in 1:dim(dat_study)[1]) {
      if (i == j) {
        varcovmat[i, j] = dat_study$var.g[i]
      } else{
        if (dat_study[i, "Experiment"] == dat_study[j, "Experiment"] &
            dat_study[i, "Level_C"] == dat_study[j, "Level_C"] &
            dat_study[i, "Trait"] == dat_study[j, "Trait"] &
            dat_study[i, "Trait.category"] == dat_study[j, "Trait.category"]) {
          varcovmat[i, j] = 1 / dat_study[i, "N_C"] + dat_study$g[i] * dat_study$g[j] /
            (dat_study[i, "N_C"] + dat_study[i, "N_X"] + dat_study[j, "N_X"])
        }
      }
    }
  }
  
  val <- eigen(varcovmat)
  for (m in 1:length(val$values)) {
    if (val$values[m] < 0) {
    studies_to_check <- append(studies_to_check, k)
    negative_eigen <- append(negative_eigen, val$values[m])
    }
  }
}
    
Neg_eigen_2 <- data.frame(study = studies_to_check, eigen = negative_eigen)

Neg_eigen_2 %>%
  kbl() %>%
  kable_material(c("striped", "hover"), full_width = F)
```

For data set 3
```{r check-var-covar_3}
studies_3 <- unique(dat_3$Study)
studies_to_check <- c()
negative_eigen <- c()

for (k in studies_3) {
  dat_study <- dat_3[which(dat_3$Study == k),]
  varcovmat = matrix(0,
                     nrow = dim(dat_study)[1],
                     ncol = dim(dat_study)[1])
  
  for (i in 1:dim(dat_study)[1]) {
    for (j in 1:dim(dat_study)[1]) {
      if (i == j) {
        varcovmat[i, j] = dat_study$var.g[i]
      } else{
        if (dat_study[i, "Experiment"] == dat_study[j, "Experiment"] &
            dat_study[i, "Level_C"] == dat_study[j, "Level_C"] &
            dat_study[i, "Trait"] == dat_study[j, "Trait"] &
            dat_study[i, "Trait.category"] == dat_study[j, "Trait.category"]) {
          varcovmat[i, j] = 1 / dat_study[i, "N_C"] + dat_study$g[i] * dat_study$g[j] /
            (dat_study[i, "N_C"] + dat_study[i, "N_X"] + dat_study[j, "N_X"])
        }
      }
    }
  }
  
  val <- eigen(varcovmat)
  for (m in 1:length(val$values)) {
    if (val$values[m] < 0) {
    studies_to_check <- append(studies_to_check, k)
    negative_eigen <- append(negative_eigen, val$values[m])
    }
  }
}
    
Neg_eigen_3 <- data.frame(study = studies_to_check, eigen = negative_eigen)

Neg_eigen_3 %>%
  kbl() %>%
  kable_material(c("striped", "hover"), full_width = F)
```

Based on these eigenvalues, we looked back at the studies/experiments and determined if they had issues with reporting N or variance. We decided to remove two such studies for data set 1 (Ashraf et al. 2017 and Shostak et al. 2015) and one such study for data set 2 & 3 (Ashraf et al. 2017). 

As we found no errors or causes of concern for other studies with barely negative eigenvalues, we are forcing the var-covar matrices to the nearest positive definite values, while keeping the diagonals (var) to the estimated values and modifying only the expected covariances.

Also for now we are excluding Experiment 121 from Civitello et al 2020 Proc. B. This study has a relatively low variance for the control of the uninfected demographic fecundity effects, which results in a negative eigenvalue.

```{r var-covar-matrix}
# exclude suspicious studies with highly negative eigenvalues
dat_1 <- dat_1[-grep("Ashraf et al. 2017 Environ Sci Pollut Res", dat_1$Study),]
dat_1 <- dat_1[-grep("Shostak et al. 2015. Journal of Parasitology", dat_1$Study),]
dat_1 <- dat_1[-grep(121, dat_1$Experiment),]

varcovmat_1 = matrix(0, nrow = dim(dat_1)[1], ncol = dim(dat_1)[1])

for (i in 1:dim(dat_1)[1]) {
  for (j in 1:dim(dat_1)[1]) {
    if (i == j) {varcovmat_1[i,j] = dat_1$var.g[i]}else{
      if (dat_1[i, "Experiment"] == dat_1[j, "Experiment"] & dat_1[i, "Level_C"] == dat_1[j, "Level_C"] & dat_1[i, "Trait"] == dat_1[j, "Trait"] & dat_1[i, "Trait.category"] == dat_1[j, "Trait.category"]) {
        varcovmat_1[i,j] = 1/dat_1[i,"N_C"] + dat_1$g[i]*dat_1$g[j]/(dat_1[i,"N_C"] + dat_1[i,"N_X"] + dat_1[j,"N_X"]) 
      }
    }
  }
}

# correct negative eigens in the few studies with large covar to var ratios
varcovmat_1_PD <- nearPD(varcovmat_1,  keepDiag = TRUE)

# exclude suspicious studies with highly negative eigenvalues
dat_2 <- dat_2[-grep("Ashraf et al. 2017 Environ Sci Pollut Res", dat_2$Study),]

varcovmat_2 = matrix(0, nrow = dim(dat_2)[1], ncol = dim(dat_2)[1])

for (i in 1:dim(dat_2)[1]) {
  for (j in 1:dim(dat_2)[1]) {
    if (i == j) {varcovmat_2[i,j] = dat_2$var.g[i]}else{
      if (dat_2[i, "Experiment"] == dat_2[j, "Experiment"] & dat_2[i, "Level_C"] == dat_2[j, "Level_C"] & dat_2[i, "Trait"] == dat_2[j, "Trait"] & dat_2[i, "Trait.category"] == dat_2[j, "Trait.category"]) {
        varcovmat_2[i,j] = 1/dat_2[i,"N_C"] + dat_2$g[i]*dat_2$g[j]/(dat_2[i,"N_C"] + dat_2[i,"N_X"] + dat_2[j,"N_X"]) 
      }
    }
  }
}

# correct negative eigens in the few studies with large covar to var ratios
varcovmat_2_PD <- nearPD(varcovmat_2,  keepDiag = TRUE)

# exclude suspicious studies with highly negative eigenvalues
dat_3 <- dat_3[-grep("Ashraf et al. 2017 Environ Sci Pollut Res", dat_3$Study),]

# Carrington et al. 2013 has extremely large sampling variance for a parasite prevalence effect. DOUBLE CHECK!!!
dat_3 <- dat_3[-grep("Carrington et al. 2013. PLOS NegTropDiseases", dat_3$Study),]

varcovmat_3 = matrix(0, nrow = dim(dat_3)[1], ncol = dim(dat_3)[1])

for (i in 1:dim(dat_3)[1]) {
  for (j in 1:dim(dat_3)[1]) {
    if (i == j) {varcovmat_3[i,j] = dat_3$var.g[i]}else{
      if (dat_3[i, "Experiment"] == dat_3[j, "Experiment"] & dat_3[i, "Level_C"] == dat_3[j, "Level_C"] & dat_3[i, "Trait"] == dat_3[j, "Trait"] & dat_3[i, "Trait.category"] == dat_3[j, "Trait.category"]) {
        varcovmat_3[i,j] = 1/dat_3[i,"N_C"] + dat_3$g[i]*dat_3$g[j]/(dat_3[i,"N_C"] + dat_3[i,"N_X"] + dat_3[j,"N_X"]) 
      }
    }
  }
}  

# correct eigenvalues in the few studies with large covar to var ratios
varcovmat_3_PD <- nearPD(varcovmat_3, keepDiag = TRUE)

# For now remove rare host types to check interaction with host
dat_3_byHost <- dat_3[which(dat_3$Host.type == "Arthropod" | dat_3$Host.type == "Mollusc" | dat_3$Host.type == "Fish" | dat_3$Host.type == "Amphibian"), ]

varcovmat_3_byHost = matrix(0, nrow = dim(dat_3_byHost)[1], ncol = dim(dat_3_byHost)[1])

for (i in 1:dim(dat_3_byHost)[1]) {
  for (j in 1:dim(dat_3_byHost)[1]) {
    if (i == j) {varcovmat_3_byHost[i,j] = dat_3_byHost$var.g[i]}else{
      if (dat_3_byHost[i, "Experiment"] == dat_3_byHost[j, "Experiment"] & dat_3_byHost[i, "Level_C"] == dat_3_byHost[j, "Level_C"] & dat_3_byHost[i, "Trait"] == dat_3_byHost[j, "Trait"] & dat_3_byHost[i, "Trait.category"] == dat_3_byHost[j, "Trait.category"]) {
        varcovmat_3_byHost[i,j] = 1/dat_3_byHost[i,"N_C"] + dat_3_byHost$g[i]*dat_3_byHost$g[j]/(dat_3_byHost[i,"N_C"] + dat_3_byHost[i,"N_X"] + dat_3_byHost[j,"N_X"]) 
      }
    }
  }
}

# correct eigenvalues in the few studies with large covar to var ratios
varcovmat_3_byHost_PD <- nearPD(varcovmat_3_byHost,  keepDiag = TRUE)
```

Create a new column with a category of vertebrate or invertebrate
```{r vertebrate/invertebrate host category, echo=FALSE}
dat_1 <- mutate(dat_1, Host.type.2 = case_when(
  Host.type == "Fish" ~ "Vertebrate",
  Host.type == "Arthropod" ~ "Invertebrate",
  Host.type == "Amphibian" ~ "Vertebrate",
  Host.type == "Mollusc" ~ "Invertebrate"))

dat_2 <- mutate(dat_2, Host.type.2 = case_when(
  Host.type == "Fish" ~ "Vertebrate",
  Host.type == "Arthropod" ~ "Invertebrate",
  Host.type == "Amphibian" ~ "Vertebrate",
  Host.type == "Mollusc" ~ "Invertebrate",
  Host.type == "Reptile" ~ "Vertebrate",
  Host.type == "Bird" ~ "Vertebrate",
  Host.type == "Mammal" ~ "Vertebrate"))

dat_3 <- mutate(dat_3, Host.type.2 = case_when( 
  Host.type == "Fish" ~ "Vertebrate",
  Host.type == "Arthropod" ~ "Invertebrate",
  Host.type == "Amphibian" ~ "Vertebrate",
  Host.type == "Mollusc" ~ "Invertebrate",
  Host.type == "Reptile" ~ "Vertebrate",
  Host.type == "Bird" ~ "Vertebrate",
  Host.type == "Mammal" ~ "Vertebrate"))

```


# Descriptive statistics
```{r descriptive-stats, echo=FALSE}
d_stats <-
  data.frame(
    stat = c(
      "Number of papers",
      "Number of effects",
      "Number of experiments",
      "Number of arthropod species",
      "Number of mollusc species",
      "Number of fish species",
      "Number of amphibian species",
      "Number of reptile species",
      "Number of bird species",
      "Number of mammal species",
      "Number of viral taxa",
      "Number of bacterial taxa",
      "Number of Multiple infection",
      "Number of fungal taxa",
      "Number of protozan taxa",
      "Number of helminth taxa",
      "Number of environment experiments",
      "Number of pollution experiments",
      "Number of resource experiments",
      "Number of resistance effects",
      "Number of prevalence effects",
      "Number of intensity effects",
      "Number of fitness effects",
      "Number of survival effects",
      "Number of fecundity effects"
    ),
    Question_1 = c(
      length(unique(dat_1$Study)),
      length(dat_1$g),
      length(unique(dat_1$Experiment)),
      length(unique(dat_1$Host[which(dat_1$Host.type == "Arthropod")])),
      length(unique(dat_1$Host[which(dat_1$Host.type == "Mollusc")])),
      length(unique(dat_1$Host[which(dat_1$Host.type == "Fish")])),
      length(unique(dat_1$Host[which(dat_1$Host.type == "Amphibian")])),
      length(unique(dat_1$Host[which(dat_1$Host.type == "Reptile")])),
      length(unique(dat_1$Host[which(dat_1$Host.type == "Bird")])),
      length(unique(dat_1$Host[which(dat_1$Host.type == "Mammal")])),
      length(unique(dat_1$Parasite[which(dat_1$Parasite.type ==
                                           "Virus")])),
      length(unique(dat_1$Parasite[which(dat_1$Parasite.type ==
                                           "Bacteria")])),
      length(unique(dat_1$Parasite[which(dat_1$Parasite.type ==
                                           "Multiple")])),
      length(unique(dat_1$Parasite[which(dat_1$Parasite.type ==
                                           "Fungus")])),
      length(unique(dat_1$Parasite[which(dat_1$Parasite.type ==
                                           "Protozoan")])),
      length(unique(dat_1$Parasite[which(dat_1$Parasite.type ==
                                           "Helminth")])),
      length(dat_1$Experiment[which(dat_1$Gradient.category ==
                                      "Environment")]),
      length(dat_1$Experiment[which(dat_1$Gradient.category ==
                                      "Pollution")]),
      length(dat_1$Experiment[which(dat_1$Gradient.category ==
                                      "Resource")]),
      length(dat_1$ID[which(dat_1$Trait.category == "Epidemiological")]),
      length(dat_1$ID[which(dat_1$Trait.type == "Prevalence")]),
      length(dat_1$ID[which(dat_1$Trait.type == "Intensity")]),
      length(dat_1$ID[grep("demographic", dat_1$Trait.category)]),
      length(dat_1$ID[which(dat_1$Trait.type == "Survivorship")]),
      length(dat_1$ID[which(dat_1$Trait.type == "Fecundity")])
    ),
    Question_2 = c(
      length(unique(dat_2$Study)),
      length(dat_2$g),
      length(unique(dat_2$Experiment)),
      length(unique(dat_2$Host[which(dat_2$Host.type == "Arthropod")])),
      length(unique(dat_2$Host[which(dat_2$Host.type == "Mollusc")])),
      length(unique(dat_2$Host[which(dat_2$Host.type == "Fish")])),
      length(unique(dat_2$Host[which(dat_2$Host.type == "Amphibian")])),
      length(unique(dat_2$Host[which(dat_2$Host.type == "Reptile")])),
      length(unique(dat_2$Host[which(dat_2$Host.type == "Bird")])),
      length(unique(dat_2$Host[which(dat_2$Host.type == "Mammal")])),
      length(unique(dat_2$Parasite[which(dat_2$Parasite.type ==
                                           "Virus")])),
      length(unique(dat_2$Parasite[which(dat_2$Parasite.type ==
                                           "Bacteria")])),
      length(unique(dat_2$Parasite[which(dat_2$Parasite.type ==
                                           "Multiple")])),
      length(unique(dat_2$Parasite[which(dat_2$Parasite.type ==
                                           "Fungus")])),
      length(unique(dat_2$Parasite[which(dat_2$Parasite.type ==
                                           "Protozoan")])),
      length(unique(dat_2$Parasite[which(dat_2$Parasite.type ==
                                           "Helminth")])),
      length(dat_2$Experiment[which(dat_2$Gradient.category ==
                                      "Environment")]),
      length(dat_2$Experiment[which(dat_2$Gradient.category ==
                                      "Pollution")]),
      length(dat_2$Experiment[which(dat_2$Gradient.category ==
                                      "Resource")]),
      length(dat_2$ID[which(dat_2$Trait.category == "Epidemiological")]),
      length(dat_2$ID[which(dat_2$Trait.type == "Prevalence")]),
      length(dat_2$ID[which(dat_2$Trait.type == "Intensity")]),
      length(dat_2$ID[grep("demographic", dat_2$Trait.category)]),
      length(dat_2$ID[which(dat_2$Trait.type == "Survivorship")]),
      length(dat_2$ID[which(dat_2$Trait.type == "Fecundity")])
    ),
    Question_3 = c(
      length(unique(dat_3$Study)),
      length(dat_3$g),
      length(unique(dat_3$Experiment)),
      length(unique(dat_3$Host[which(dat_3$Host.type == "Arthropod")])),
      length(unique(dat_3$Host[which(dat_3$Host.type == "Mollusc")])),
      length(unique(dat_3$Host[which(dat_3$Host.type == "Fish")])),
      length(unique(dat_3$Host[which(dat_3$Host.type == "Amphibian")])),
      length(unique(dat_3$Host[which(dat_3$Host.type == "Reptile")])),
      length(unique(dat_3$Host[which(dat_3$Host.type == "Bird")])),
      length(unique(dat_3$Host[which(dat_3$Host.type == "Mammal")])),
      length(unique(dat_3$Parasite[which(dat_3$Parasite.type ==
                                           "Virus")])),
      length(unique(dat_3$Parasite[which(dat_3$Parasite.type ==
                                           "Bacteria")])),
      length(unique(dat_3$Parasite[which(dat_3$Parasite.type ==
                                           "Multiple")])),
      length(unique(dat_3$Parasite[which(dat_3$Parasite.type ==
                                           "Fungus")])),
      length(unique(dat_3$Parasite[which(dat_3$Parasite.type ==
                                           "Protozoan")])),
      length(unique(dat_3$Parasite[which(dat_3$Parasite.type ==
                                           "Helminth")])),
      length(dat_3$Experiment[which(dat_3$Gradient.category ==
                                      "Environment")]),
      length(dat_3$Experiment[which(dat_3$Gradient.category ==
                                      "Pollution")]),
      length(dat_3$Experiment[which(dat_3$Gradient.category ==
                                      "Resource")]),
      length(dat_3$ID[which(dat_3$Trait.category == "Epidemiological")]),
      length(dat_3$ID[which(dat_3$Trait.type == "Prevalence")]),
      length(dat_3$ID[which(dat_3$Trait.type == "Intensity")]),
      length(dat_3$ID[grep("demographic", dat_3$Trait.category)]),
      length(dat_3$ID[which(dat_3$Trait.type == "Survivorship")]),
      length(dat_3$ID[which(dat_3$Trait.type == "Fecundity")])
    )
  )

d_stats %>%
  kbl() %>%
  kable_material(c("striped", "hover"), full_width = F)
```


# Preliminary analysis

## Question 1 - Are fitness traits negatively affected by stressors
### Null model
```{r Q1-null}
Q1m1 <-
  rma.mv(
    g ~ 1,
    V = varcovmat_1_PD$mat,
    random = list(  ~ 1 |Experiment/ID, ~ 1 | Parasite),
    data = dat_1,
    method = "REML"
  )

summary(Q1m1)
```


### Effects in uninfected vs infected hosts, do they differe between survival vs fecundity?
```{r Q1-single-fixed}
Q1m2 <-
  rma.mv(
    g ~ Trait.category:Trait.type -1,
    V = varcovmat_1_PD$mat,
    random = list( ~ 1 | as.factor(ID), ~ 1 | as.factor(Experiment), ~ 1 | Parasite),
    data = dat_1,
    method = "REML"
  )

summary(Q1m2)
forest.default(x= Q1m2$beta, sei =  Q1m2$se, ci.lb =  Q1m2$ci.lb, ci.ub =  Q1m2$ci.ub, 
                annotate=TRUE, showweights=T, header=F,
                slab = c("Infected fecundity", "Uninfected fecundity", 
                          "Infected survivorship", "Uninfected survivorship"))
```

### Effects in uninfected vs infected hosts, do they differe between survival vs fecundity, once we account for type of stress
```{r Q1-single-fixed}
Q1m3 <-
  rma.mv(
    g ~ Trait.category : Trait.type : Gradient.category -1,
    V = varcovmat_1_PD$mat,
    random = list( ~ 1 | Experiment/ID, ~ 1 | Parasite),
    data = dat_1,
    method = "REML"
  )

summary(Q1m3)
forest.default(x= Q1m3$beta, sei =  Q1m3$se, ci.lb =  Q1m3$ci.lb, ci.ub =  Q1m3$ci.ub, 
                annotate=TRUE, showweights=T, header=F,
                slab = c("Environment infected fecundity", "Environment uninfected fecundity", "Environment infected survivorship", "Environment uninfected survivorship", "Pollution infected fecundity", "Pollution uninfected fecundity", "Pollution infected survivorship", "Pollution uninfected survivorship", "Resource infected fecundity", "Resource uninfected fecundity", "Resource infected survivorship", "Resource uninfected survivorship"))
```

### Effects in  survival vs fecundity measures of fitness under different types of stress, combining infected and uninfected
```{r Q1-single-fixed}
Q1m4 <-
  rma.mv(
    g ~ Trait.type:Gradient.category -1 ,
    V = varcovmat_1_PD$mat,
    random = list( ~ 1 | Experiment/ID, ~ 1 | Parasite),
    data = dat_1,
    method = "REML"
  )

summary(Q1m4)
forest.default(x= Q1m4$beta, sei =  Q1m4$se, ci.lb =  Q1m4$ci.lb, ci.ub =  Q1m4$ci.ub, 
                annotate=TRUE, showweights=T, header=F,
                slab = c("Environment fecundity", "Environment survivorship", "Pollution  fecundity", "Pollution  survivorship", "Resource  fecundity", "Resource  survivorship"))
```

### Trait category x Host type interaction
```{r Q1-Trait-x-Host}
Q1m5 <-
  rma.mv(
    g ~ Trait.category:Host.type.2 -1,
    V = varcovmat_1_PD$mat,
    random = list( ~ 1 | ID, ~ 1 | Experiment, ~ 1 | Parasite),
    data = dat_1,
    method = "REML"
  )

summary(Q1m5)
forest.default(x= Q1m5$beta, sei =  Q1m5$se, ci.lb =  Q1m5$ci.lb, ci.ub =  Q1m5$ci.ub, 
                annotate=TRUE, showweights=T, header=F,
                slab = c("Invertebrate infected", "Invertebrate uninfected", "Vertebrate infected", "Vertebrate  uninfected"))
```

### Trait category x Stress type x Host type interaction
```{r Q1-Trait-x-Host}
Q1m6 <-
  rma.mv(
    g ~ Trait.category:Gradient.category:Host.type.2 -1,
    V = varcovmat_1_PD$mat,
    random = list( ~ 1 | ID, ~ 1 | Experiment, ~ 1 | Parasite),
    data = dat_1,
    method = "REML"
  )

summary(Q1m6)
forest.default(x= Q1m6$beta, sei =  Q1m6$se, ci.lb =  Q1m6$ci.lb, ci.ub =  Q1m6$ci.ub, 
                annotate=TRUE, showweights=T, header=F,
                slab = c("Environment invertebrate infected", "Environment invertebrate uninfected", "Pollution invertebrate infected", "Pollution invertebrate uninfected", "Resource invertebrate infected", "Resource invertebrate uninfected", "Environment vertebrate infected", "Environment vertebrate uninfected", "Pollution vertebrate infected", "Pollution vertebrate uninfected", "Resource vertebrate infected", "Resource vertebrate uninfected"))
```

### Trait type x Stress type x Host type interaction
This model should not be included in the model selection, because it is missing categories (we don't have data of fecundity from vertebrates)
```{r Q1-Trait-x-Host}
Q1m7 <-
  rma.mv(
    g ~ Trait.type:Gradient.category:Host.type.2 -1,
    V = varcovmat_1_PD$mat,
    random = list( ~ 1 | ID, ~ 1 | Experiment, ~ 1 | Parasite),
    data = dat_1,
    method = "REML"
  )

summary(Q1m7)
forest.default(x= Q1m7$beta, sei =  Q1m7$se, ci.lb =  Q1m7$ci.lb, ci.ub =  Q1m7$ci.ub, 
                annotate=TRUE, showweights=T, header=F,
                slab = c("Environment invertebrate fecundity", "Environment invertebrate survivorship", "Pollution invertebrate fecundity", "Pollution invertebrate survivorship", "Resource invertebrate fecundity", "Resource invertebrate survivorship", "Environment vertebrate survivorship", "Pollution vertebrate survivorship", "Resource vertebrate survivorship"))
```

### Stress type x Host type interaction, combining infected and uninfected
```{r Q1-Trait-x-Host}
Q1m8 <-
  rma.mv(
    g ~ Gradient.category:Host.type.2 -1,
    V = varcovmat_1_PD$mat,
    random = list( ~ 1 | ID, ~ 1 | Experiment, ~ 1 | Parasite),
    data = dat_1,
    method = "REML"
  )

summary(Q1m8)
forest.default(x= Q1m8$beta, sei =  Q1m8$se, ci.lb =  Q1m8$ci.lb, ci.ub =  Q1m8$ci.ub, 
                annotate=TRUE, showweights=T, header=F,
                slab = c("Environment invertebrate", "Pollution invertebrate", "Resource invertebrate", "Environment vertebrate", "Pollution vertebrate", "Resource vertebrate "))
```

### Model selection
#### Using MuMIn package
```{r}
# First, we evaluate some code that generates helper functions needed so that metafor and MuMIn could interact as necessary
eval(metafor:::.MuMIn)

# Now, we take the full model and fit the rest of the models and examine those models whose AICc value is no more than 10 units away from that of the best model
full_model.1 <- rma.mv(
    g,
    V = varcovmat_1_PD$mat,
    mods = ~ Trait.category*Trait.type*Gradient.category, 
    random = list( ~ 1 | ID, ~ 1 | Experiment, ~ 1 | Parasite),
    data = dat_1,
    method = "ML"
  )

model_selection.1 <- dredge(full_model.1, trace = 2) 
subset(model_selection.1, delta <= 10, recalc.weights = FALSE)
# --> I am getting different weights (and therefore different unconditional estimates) than with the glmulti method

# Multimodel inference
average.model.1 <- model.avg(model_selection.1, revised.var=FALSE)
summary(average.model.1) #Set revised.var = F to get same results as we would with the glmulti package

# relative importance values for the predictors can be obtained with:
importance(model_selection.1)

# Best model Q1m4
```

## Question 2 - Are resistance traits negatively affected by stressors
### Null model
```{r Q2-null}
Q2m1 <-
  rma.mv(
    g ~ 1,
    V = varcovmat_2_PD$mat,
    random = list( ~ 1 | as.factor(ID), ~ 1 | as.factor(Experiment), ~ 1 | Parasite),
    data = dat_2,
    method = "REML"
  )

summary(Q2m1)
```

### Does resistance decrease with stress, does it differe between prevalence vs intensity?
```{r Q2-single-fixed}
Q2m2 <-
  rma.mv(
    g ~ Trait.type -1 ,
    V = varcovmat_2_PD$mat,
    random = list( ~ 1 | as.factor(ID), ~ 1 | as.factor(Experiment), ~ 1 | Parasite),
    data = dat_2,
    method = "REML"
  )

summary(Q2m2)
forest.default(x= Q2m2$beta, sei =  Q2m2$se, ci.lb =  Q2m2$ci.lb, ci.ub =  Q2m2$ci.ub, 
                annotate=TRUE, showweights=T, header=F,
                slab = c("Intensity", "Prevalence"))
```

### Do they differe between prevalence vs intensity, once we account for type of stress
```{r Q2-single-fixed}
Q2m3 <-
  rma.mv(
    g ~ Trait.type : Gradient.category -1,
    V = varcovmat_2_PD$mat,
    random = list( ~ 1 |Experiment/ID, ~ 1 | Parasite),
    data = dat_2,
    method = "REML"
  )

summary(Q2m3)
forest.default(x= Q2m3$beta, sei =  Q2m3$se, ci.lb =  Q2m3$ci.lb, ci.ub =  Q2m3$ci.ub, 
                annotate=TRUE, showweights=T, header=F,
                slab = c("Environment intensity", "Environment prevalence", "Pollution intensity", "Pollution prevalence", "Resource intensity", "Resource prevalence"))
```


### Trait type x Host type interaction
```{r Q2-Trait-x-Host}
Q2m4 <-
  rma.mv(
    g ~ Trait.type:Host.type.2:Gradient.category -1,
    V = varcovmat_2_PD$mat,
    random = list( ~ 1 | ID, ~ 1 | Experiment, ~ 1 | Parasite),
    data = dat_2,
    method = "REML"
  )

summary(Q2m4)
forest.default(x= Q2m4$beta, sei =  Q2m4$se, ci.lb =  Q2m4$ci.lb, ci.ub =  Q2m4$ci.ub, 
                annotate=TRUE, showweights=T, header=F,
                slab = c("Environment invertebrate intensity", "Environment invertebrate prevalence", "Environment vertebrate intensity", "Environment vertebrate prevalence", "Pollution invertebrate intensity", "Pollution invertebrate prevalence", "Pollution vertebrate intensity", "Pollution vertebrate prevalence", "Resource invertebrate intensity", "Resource invertebrate prevalence", "Resource vertebrate intensity", "Resource vertebrate prevalence"))
```
### Does the response depends on stress type 
```{r Q2-Stress-Type}
Q2m5 <-
  rma.mv(
    g ~ Gradient.category -1,
    V = varcovmat_2_PD$mat,
    random = list( ~ 1 | ID, ~ 1 | Experiment, ~ 1 | Parasite),
    data = dat_2,
    method = "REML"
  )

summary(Q2m5)
forest.default(x= Q2m5$beta, sei =  Q2m5$se, ci.lb =  Q2m5$ci.lb, ci.ub =  Q2m5$ci.ub, 
                annotate=TRUE, showweights=T, header=F,
                slab = c("Environment", "Pollution", "Resource"))
```
### Model selection
This section is a work in progress...

#### Using glmulti package
```{r}
# Define a function that takes a model formula and a dataset as input and then fits a random/mixed-effects meta-regression model to the given data using maximum likelihood estimation (to compare models we have to use ML, and then we can re-fit with REML)
rma.glmulti <- function(formula, data, V, random, ...){
   rma.mv(formula, V = V, data = data, random = random, method="ML", ...) 
}

# Fit all models
all.models.2 <- glmulti(g ~ Trait.type*Gradient.category, 
               data=dat_2,
               V = varcovmat_2_PD$mat, 
               random = list( ~ 1 | ID, ~ 1 | Experiment, ~ 1 | Parasite),
               level = 2, fitfunction = rma.glmulti, crit="aicc", marginality = TRUE)

# --> Tried "intercept = F", to exclude the intercept from  the candidate models, and got "Error in x[, ii] : subscript out of bounds"

# summary
print(all.models.2)

# plot of AICc values for the models. The line separates the models that are less or more that 2 units away from the best model
plot(all.models.2)

# Top models
top.models.2 <- weightable(all.models.2)
top.models.2 <- top.models.2[top.models.2$aicc <= min(top.models.2$aicc) + 2,]
top.models.2

# Variable importance plot across all the models
plot(all.models.2, type="s")
# --> none of them reach >= 0.8

#Multimodel inference - make inferences about various predictors across all possible models relative to their weights

#To get the glmulti package to handle rma.uni objects, we must register a getfit method for such objects
eval(metafor:::.glmulti)

coef(all.models.2)

# Organized table, getting the unconditional estimates
mmi.2 <- as.data.frame(coef(all.models.2))
mmi.2 <- data.frame(Estimate=mmi.2$Est, SE=sqrt(mmi.2$Uncond), Importance=mmi.2$Importance, row.names=row.names(mmi.2))
mmi.2$z <- mmi.2$Estimate / mmi.2$SE
mmi.2$p <- 2*pnorm(abs(mmi.2$z), lower.tail=FALSE)
names(mmi.2) <- c("Estimate", "Std. Error", "Importance", "z value", "Pr(>|z|)")
mmi.2$ci.lb <- mmi.2[[1]] - qnorm(.975) * mmi.2[[2]]
mmi.2$ci.ub <- mmi.2[[1]] + qnorm(.975) * mmi.2[[2]]
mmi.2 <- mmi.2[order(mmi.2$Importance, decreasing=TRUE), c(1,2,4:7,3)]
round(mmi.2, 4)

#Multimodel predictions
# --> Note: how (and when) can we re-fit the model with REML?
#           how do we deal with interactions?

# x <- data.frame("Gradient.category" = c("Environment", "Pollution", "Resource", "Environment", "Pollution", "Resource"), "Trait.type" = c("Intensity", "Intensity", "Intensity", "Prevalence", "Prevalence", "Prevalence"))
# predict.glmulti(res, varweighting = "Buckland", newdata = x, icmethod = "Lukacs", se.fit = TRUE)

preds <- list()

x <- c("Gradient.categoryPollution" = T, "Gradient.categoryResource" = F, "Trait.typePrevalence" = T, "Gradient.categoryEnvironment:Trait.typeIntensity" = F, "Gradient.categoryPollution:Trait.typeIntensity" = F, 
"Gradient.categoryResource:Trait.typeIntensity" = F, "Gradient.categoryEnvironment:Trait.typePrevalence" = F,
"Gradient.categoryPollution:Trait.typePrevalence" = F, 
"Gradient.categoryResource:Trait.typePrevalence" = F,
"Trait.typePrevalence:Gradient.categoryResource" = F, 
"Trait.typePrevalence:Gradient.categoryPollution" = F,
"Trait.typePrevalence:Gradient.categoryEnvironment" = F,
"Trait.typeIntensity:Gradient.categoryResource" = F,
"Trait.typeIntensity:Gradient.categoryPollution" = F,
"Trait.typeIntensity:Gradient.categoryEnvironment" = F)

#Missing Gradient.categoryEnvironment and Trait.typeIntensity
 
for (j in 1:all.models.2@nbmods) {
   model <- all.models.2@objects[[j]]
   vars <- names(coef(model))[-1]
   if (length(vars) == 0) {
      preds[[j]] <- predict(model)
   } else {
      preds[[j]] <- predict(model, newmods=x[vars])
   }
}

weights <- weightable(all.models.2)$weights
yhat <- sum(weights * sapply(preds, function(x) x$pred))
round(yhat, 3)
se <- sqrt(sum(weights * sapply(preds, function(x) x$se^2 + (x$pred - yhat)^2)))
round(yhat + c(-1,1)*qnorm(.975)*se, 3)

```

#### Using MuMIn package
```{r}
# Now, we take the full model and fit the rest of the models and examine those models whose AICc value is no more than 10 units away from that of the best model
full_model.2 <- rma.mv(
    g,
    V = varcovmat_2_PD$mat,
    mods = ~ Trait.type * Gradient.category, 
    random = list( ~ 1 | ID, ~ 1 | Experiment, ~ 1 | Parasite),
    data = dat_2,
    method = "ML"
  )

model_selection.2 <- dredge(full_model.2, trace = 2) 
subset(model_selection.2, delta <= 10, recalc.weights = FALSE)
# --> I am getting different weights (and therefore different unconditional estimates) than with the glmulti method

# Multimodel inference
average.model.2 <- model.avg(model_selection.2, revised.var=FALSE)
summary(average.model.2) #Set revised.var = F to get same results as we would with the glmulti package

# relative importance values for the predictors can be obtained with:
importance(model_selection.2)

#How do we get model predictions? - Still not working
predict(average.model.2, se.fit = T)
predict(average.model.2, newdata = data.frame(c("Gradient.categoryPollution" = T, "Gradient.categoryResource" = F, "Trait.typePrevalence" = T, "Gradient.categoryPollution:Trait.typePrevalence" = F, 
"Gradient.categoryResource:Trait.typePrevalence" = F)), se.fit = T)
```


## Question 3 - Do resistance and fitness traits have different sensitivity to stressors
### Null model
```{r Q3-null}
Q3m1 <-
  rma.mv(
    g ~ 1,
    V = varcovmat_3_PD$mat,
    random = list( ~ 1 | ID, ~ 1 | Experiment, ~ 1 | Parasite),
    data = dat_3,
    method = "REML"
  )

summary(Q3m1)

```

### Single fixed effect
```{r Q3-single-fixed}
#note that this is WITH absolute value --> which one is more sensible. Both are significant.
Q3m2.1 <-
  rma.mv(
   abs(g) ~ Trait.category -1,
    V = varcovmat_3_PD$mat,
    random = list( ~ 1 | ID, ~ 1 | Experiment, ~ 1 | Parasite),
    data = dat_3,
    method = "REML"
  )

summary(Q3m2.1)
forest.default(x= Q3m2.1$beta, sei =  Q3m2.1$se, ci.lb =  Q3m2.1$ci.lb, ci.ub =  Q3m2.1$ci.ub, 
                annotate=TRUE, showweights=T, header=F,
                slab = c("epidemiological", "Fitness"))
```
```{r Q3-single-fixed}
#note that this is WITHOUT absolute value --> In here, resistance is not significant. Which means that the responses to stress are all over the place, while for the fitness trait is always a negative effect. Could be explained by dying individuals when getting infected, which might look like an increase in resistance (??)

Q3m2.2 <-
  rma.mv(
   g ~ Trait.category -1,
    V = varcovmat_3_PD$mat,
    random = list( ~ 1 | ID, ~ 1 | Experiment, ~ 1 | Parasite),
    data = dat_3,
    method = "REML"
  )

summary(Q3m2.2)
forest.default(x= Q3m2.2$beta, sei =  Q3m2.2$se, ci.lb =  Q3m2.2$ci.lb, ci.ub =  Q3m2.2$ci.ub, 
                annotate=TRUE, showweights=T, header=F,
                slab = c("Epidemiological", "Fitness"))
```

### Trait category x Gradient category interaction
```{r Q3-Trait-x-Gradient}
Q3m3 <-
  rma.mv(
    g ~ Trait.category:Gradient.category -1,
    V = varcovmat_3_PD$mat,
    random = list( ~ 1 | ID, ~ 1 | Experiment, ~ 1 | Parasite),
    data = dat_3,
    method = "REML"
  )

summary(Q3m3)
forest.default(x= Q3m3$beta, sei =  Q3m3$se, ci.lb =  Q3m3$ci.lb, ci.ub =  Q3m3$ci.ub, 
                annotate=TRUE, showweights=T, header=F,
                slab = c("Environment epidemiological", "Environment fitness", "Pollution epidemiological", "Pollution fitness", "Resource epidemiological", "Resource fitness"))
```

### Trait type x Gradient category interaction
```{r Q3-Trait-x-Gradient}
Q3m4 <-
  rma.mv(
    g ~ Trait.type:Gradient.category -1,
    V = varcovmat_3_PD$mat,
    random = list( ~ 1 | Experiment/ID, ~ 1 | Parasite),
    data = dat_3,
    method = "REML",
)

summary(Q3m4)
forest.default(x= Q3m4$beta, sei =  Q3m4$se, ci.lb =  Q3m4$ci.lb, ci.ub =  Q3m4$ci.ub, 
                annotate=TRUE, showweights=T, header=F,
                slab = c("Environment fecundity", "Environment intensity", "Environment prevalence", "Environment survivorship", "Pollution fecundity", "Pollution intensity", "Pollution prevalence", "Pollution survivorship", "Resource fecundity", "Resource intensity", "Resource prevalence", "Resource survivorship"))
```

### Trait category x Host type interaction
```{r Q3-Trait-x-Host}
Q3m5 <-
  rma.mv(
    g ~ Trait.category:Host.type.2 -1,
    V = varcovmat_3_PD$mat,
    random = list( ~ 1 | ID, ~ 1 | Experiment, ~ 1 | Parasite),
    data = dat_3,
    method = "REML",
)

summary(Q3m5)
forest.default(x= Q3m5$beta, sei =  Q3m5$se, ci.lb =  Q3m5$ci.lb, ci.ub =  Q3m5$ci.ub, 
                annotate=TRUE, showweights=T, header=F,
                slab = c("Resistance invertebrate", "Fitness invertebrate", "Resistance vertebrate", "Fitness vertebrate"))
```

### Trait category x Invert/Vert x Stressor type
```{r Q3-Trait-x-Host}
Q3m6 <-
  rma.mv(
    g ~ Trait.category:Host.type.2:Gradient.category -1,
    V = varcovmat_3_PD$mat,
    random = list( ~ 1 | ID, ~ 1 | Experiment, ~ 1 | Parasite),
    data = dat_3,
    method = "REML",
)

summary(Q3m6)
forest.default(x= Q3m6$beta, sei =  Q3m6$se, ci.lb =  Q3m6$ci.lb, ci.ub =  Q3m6$ci.ub, 
                annotate=TRUE, showweights=T, header=F,
                slab = c("Environment invertebrate epidemiological", "Environment invertebrate fitness", "Environment vertebrate epidemiological", "Environment vertebrate fitness", "Pollution invertebrate epidemiological", "Pollution invertebrate fitness", "Pollution vertebrate epidemiological", "Pollution vertebrate fitness", "Resource invertebrate epidemiological", "Resource invertebrate fitness", "Resource vertebrate epidemiological", "Resource vertebrate fitness"))

```

### Trait type x Invert/Vert x stressor type
```{r Q3-Trait-x-Host}
Q3m7 <-
  rma.mv(
    g ~ Trait.type:Host.type.2:Gradient.category -1,
    V = varcovmat_3_PD$mat,
    random = list( ~ 1 | ID, ~ 1 | Experiment, ~ 1 | Parasite),
    data = dat_3,
    method = "REML",
)

summary(Q3m7)
forest.default(x= Q3m7$beta, sei =  Q3m7$se, ci.lb =  Q3m7$ci.lb, ci.ub =  Q3m7$ci.ub, 
                annotate=TRUE, showweights=T, header=F,
                slab = c("Environment invertebrate fecundity", "Environment invertebrate intensity", "Environment invertebrate prevalence", "Environment invertebrate survivorship", "Environment vertebrate intensity", "Environment vertebrate prevalence", "Environment vertebrate survivorship", "Pollution invertebrate fecundity", "Pollution invertebrate intensity", "Pollution invertebrate prevalence", "Pollution invertebrate survivorship", "Pollution vertebrate intensity", "Pollution vertebrate prevalence", "Pollution vertebrate survivorship", "Resource invertebrate fecundity", "Resource invertebrate intensity", "Resource invertebrate prevalence", "Resource invertebrate survivorship", "Resource vertebrate intensity", "Resource vertebrate prevalence", "Resource vertebrate survivorship"))

```

### Model selection
#### Using MuMIn package
```{r}
full_model.3 <- rma.mv(
    g,
    V = varcovmat_3_PD$mat,
    mods = ~ Trait.type*Gradient.category, 
    random = list( ~ 1 | ID, ~ 1 | Experiment, ~ 1 | Parasite),
    data = dat_3,
    method = "ML"
  )
# Define if using Trait.type or Trait.category
model_selection.3 <- dredge(full_model.3, trace = 2) 
subset(model_selection.3, delta <= 10, recalc.weights = FALSE)
# --> I am getting different weights (and therefore different unconditional estimates) than with the glmulti method

# Multimodel inference
average.model.3 <- model.avg(model_selection.3, revised.var=FALSE)
summary(average.model.3) #Set revised.var = F to get same results as we would with the glmulti package

# relative importance values for the predictors can be obtained with:
importance(model_selection.3)

#Best model Q3m4
```

# Publication bias

The most common way to assess publication bias is to test whether studies with large standard errors/low precision/small sample size tend to have larger effect sizes. Small studies have larger SEs, and because they are small, they are more likely to get published only if results are significant. In contrast, large studies, which have higher precision and have costed more money, are more likely to get published either way. For this reason small studies can cause publication bias. 

### Funnel plots
A very common way to assess publication bias is therefore to check whether studies with low precision generally have larger effects. The typical way to visualize this is through a funnel plot with all effects. Now, this is not entirely correct for us, as the effects are not independent. Nakagawa et al 2021 (Methods Ecol Evol) suggest using conditional residuals for funnel plots. These residual plots take some of the non-independence into account, but are not perfect, as they assume: 1)sampling SE(sei) does not covary with moderators in meta-regression, 2) sampling SE is the same as the SE of the residuals and 3) sampling variances are not correlated. This last one is most certainly incorrect.

Estimation of conditional residuals is not available to rma.mv objects (which can include a sampling variance-covariance matrix). The available alternative is to ignore this covariation in sampling variance for the assessment of publication bias. This is a caveat that should be acknowledgend in the paper, yet I think it is a useful visualization nonetheless. To do these funnel plots I re-fit the best model as rma.uni for each data set. While not ideal, this the best current alternative suggested by Nakagawa et al 2021 (Methods Ecol Evol)

**Note** I checked some previous studies that use multilevel regression models as we do here, to see what they do about publication bias. Sauer et al 2020 Ecology (101) acknowledged this problem and decided not to even show funnel plots: "Because of the complex non independence among effect sizes within a study (e.g., some studies had multiple effect sizes), we did not use funnel plots or rank correlation tests to assess publication bias (Lau et al. 2006, Civitello et al.2015)." Others (Yates et al. 2019 Environmental DNA (1); Salerno et al 2021 Global Change Biology (27); Brlik et al. 2020 JAE) acknowledge this problem, but still show funnel plots warning caution. They also conduct a version of Eggerts test somewhat in the same spirit as I do below.

#### Data set 1: fitness effects of environmental stress
For now, I'm using the best model for Question 1:Q1m7
```{r dat_1-funnel}
# re fit as a mixed effects model with rma.uni
# any fitting method other than "FE" works for a random/mixed effects model
Q1m7.uni <-
  rma.uni(
    g ~ Trait.type:Gradient.category:Host.type.2 + Experiment + ID ,
    vi = dat_1$var.g,
    data = dat_1,
    method = "REML"
  )

# get conditional residuals (residuals for each effect)
Q1res <- rstandard.rma.uni(Q1m7.uni, type = "conditional")

# create colour palette as  gradient of effect size g
nlvl <- length(dat_1$g)
pal <- colorRampPalette(brewer.pal(9, "YlOrRd"))(nlvl)
pal <-pal[order(match(pal,dat_1$g))]

# plot funnel with conditional residuals
metafor::funnel(Q1res$resid, sqrt(dat_1$var.g), level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), col=pal, back = "white", refline=0, legend=TRUE, xlab = "Conditional residual", ylab = "Standard error")
```

We can see that as SE increases (e.g ~ 1) all studies report large negative fitness effects of environmental stress (large negative residuals). In other words, it seems that very small studies tend to find that environmental stress is more negative to fitness!

#### Data set 2: resistance effects of environmental stress
For now, I'm using the best model for Question 2: Q2m4
```{r dat_2-funnel}
# re fit as a mixed effects model with rma.uni
# any fitting method other than "FE" works for a random/mixed effects model
Q2m4.uni <-
  rma.uni(
    g ~ Trait.type:Gradient.category:Host.type.2 + Experiment + ID ,
    vi = dat_2$var.g,
    data = dat_2,
    method = "REML"
  )

# get conditional residuals (residuals for each effect)
Q2res <- rstandard.rma.uni(Q2m4.uni, type = "conditional")

# create colour palette as  gradient of effect size g
nlvl <- length(dat_2$g)
pal <- colorRampPalette(brewer.pal(9, "YlOrRd"))(nlvl)
pal <-pal[order(match(pal,dat_2$g))]

# plot funnel with all conditional residuals
metafor::funnel(Q2res$resid, sqrt(dat_2$var.g), level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), col=pal, back = "white", refline=0, legend=TRUE, xlab = "Conditional residual", ylab = "Standard error")
```

There is an outlier effect with extremely low precision. Amanda may want to double check this point. To be able to visualize the bulk of the data I plot the rest of the effects below.

```{r dat_2-funnel-no-outlier}
out <- grep(x = dat_2$var.g, pattern = max(dat_2$var.g))

pal <- pal[-out]

metafor::funnel(Q2res$resid[-out], sqrt(dat_2$var.g[-out]), level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), col=pal, back = "white", refline=0, legend=TRUE, xlab = "Conditional residual", ylab = "Standard error")
```

Again, it seems that studies with low precision (sei > 1) tend to have disproportionally large effects. In this case, publication bias would result in over estimation of infection intensity under stressful conditions. 

#### Data set 3: resistance and fitness effects of environmental stress
```{r dat_3-funnel}
# re fit as a mixed effects model with rma.uni
# any fitting method other than "FE" works for a random/mixed effects model
Q3m7.uni <-
  rma.uni(
    g ~ Trait.type:Gradient.category:Host.type.2 + Experiment + ID,
    vi = dat_3$var.g,
    data = dat_3,
    method = "REML"
  )

# get conditional residuals (residuals for each effect)
Q3res <- rstandard.rma.uni(Q3m7.uni, type = "conditional")

# create colour palette as  gradient of effect size g
nlvl <- length(dat_3$g)
pal <- colorRampPalette(brewer.pal(9, "YlOrRd"))(nlvl)
pal <-pal[order(match(pal,dat_3$g))]

# plot funnel with all conditional residuals
metafor::funnel(Q3res$resid, sqrt(dat_3$var.g), level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), col=pal, back = "white", refline=0, legend=TRUE, xlab = "Conditional residual", ylab = "Standard error")
```
Because this data set includes both the fitness and resistance effects it's not surprising that small studies are distributed on both sides of the funnel. This does not mean absence of publication bias as negative effects for fitness and positive effects for (-) resistance both would suggest that small studies tend to inflate the negative consequences of environmental stress. For the publication, I suggest we use -g for resistance effects (currently g = infection intensity or prevalence) and/or we use funnel plots for only datasets 1 and 2.

### Two step approach to test for publication bias
Following Nakagawa et al. (2021) we conduct a two-step approach to test for publication bias, that can be thought of as a modified Egger's test for multilevel meta analysis. First we include the effect sizes' standard error as the only moderator and including the same random effect structure as in our meta-analysis. If the slope of this moderator is significant, the estimated intercept provides a downwardly biased estimate of the overall meta-analytic effect (Doucouliagos 2012, 2014). In this case, we run a second meta regression with only the effect sizes' variances as moderators. The intercept in this case may still be biased, but less so.

#### Data set 1: fitness effects of environmental stress
For now, I'm using the best model for Question 1:Q1m7

```{r dat_1-twostep-1}
# get standard errors of effect sizes 
dat_1$sei <- sqrt(dat_1$var.g)

# meta regression with SE
rma.mv(g ~ 1 + sei,
    V = varcovmat_1_PD$mat,
    random = list( ~ 1 | Experiment, ~ 1 | ID),
    data = dat_1,
    method = "REML")
```

The slope is negative and significant. Small studies report large and negative effect sizes!! Would this be the case without accounting for covariance in sampling errors?

```{r dat_1-twostep-1-RVE}
rma.mv(g ~ 1 + sei,
    V = var.g,
    random = list( ~ 1 | Experiment, ~ 1 | ID),
    data = dat_1,
    method = "REML")
```

The results are qualitatively similar and both suggest bias. Because the slope of sei is significant, variance is a better moderator according to (Doucouliagos 2012, 2014). Here, I show for models with and without sampling error covariance within experiments

```{r dat_1-twostep-2}
rma.mv(g ~ 1 + var.g,
    V = varcovmat_1_PD$mat,
    random = list( ~ 1 | Experiment, ~ 1 | ID),
    data = dat_1,
    method = "REML")

rma.mv(g ~ 1 + var.g,
    V = var.g,
    random = list( ~ 1 | Experiment, ~ 1 | ID),
    data = dat_1,
    method = "REML")
```

Again, precision effects are similar and significant, suggesting publication bias. Thus, to "account" for potential effects of publication bias, we need to include variance as a moderator.

Including the variance-covariance error matrix within studies
```{r dat_1-PB-adjusted}
Q1m7.B <-
  rma.mv(
    g ~ Trait.type:Gradient.category:Host.type.2 + var.g -1,
    V = varcovmat_1_PD$mat,
    random = list( ~ 1 | ID, ~ 1 | Experiment),
    data = dat_1,
    method = "REML"
  )

summary(Q1m7.B)

forest.default(x= Q1m7.B$beta, sei =  Q1m7.B$se, ci.lb = Q1m7.B$ci.lb, ci.ub = Q1m7.B$ci.ub, 
                annotate=TRUE, showweights=T, header=F, 
                slab = c("variance","Environment invertebrate fecundity", "Environment invertebrate survivorship", "Pollution invertebrate fecundity", "Pollution invertebrate survivorship", "Resource invertebrate fecundity", "Resource invertebrate survivorship", "Environment vertebrate survivorship", "Pollution vertebrate survivorship", "Resource vertebrate survivorship"))
```

Using the robust variance estimator instead of the variance-covariance error matrix within studies
```{r dat_1-PB-adjusted-RVE}
Q1m7.B.R <-
  rma.mv(
    g ~ Trait.type:Gradient.category:Host.type.2 + var.g -1,
    V = var.g,
    random = list( ~ 1 | ID, ~ 1 | Experiment),
    data = dat_1,
    method = "REML"
  )

Q1m7.B.R_est <- clubSandwich::coef_test(Q1m7.B.R, vcov = "CR2")

forest.default(x= Q1m7.B.R_est$beta, sei =  Q1m7.B.R_est$SE, ci.lb = (Q1m7.B.R_est$beta - 1.96 * Q1m7.B.R_est$SE) , ci.ub =(Q1m7.B.R_est$beta + 1.96 * Q1m7.B.R_est$SE), 
                annotate=TRUE, showweights=T, header=F, 
                slab = c("variance","Environment invertebrate fecundity", "Environment invertebrate survivorship", "Pollution invertebrate fecundity", "Pollution invertebrate survivorship", "Resource invertebrate fecundity", "Resource invertebrate survivorship", "Environment vertebrate survivorship", "Pollution vertebrate survivorship", "Resource vertebrate survivorship"))
```

#### Data set 2: resistance effects of environmental stress

For now, I'm using the best model for Question 2: Q2m4

```{r dat_2-twostep-1}
# get standard errors of effect sizes 
dat_2$sei <- sqrt(dat_2$var.g)

# meta regression with SE
rma.mv(g ~ 1 + sei,
    V = varcovmat_2_PD$mat,
    random = list( ~ 1 | Experiment, ~ 1 | ID),
    data = dat_2,
    method = "REML")
```

Without var-covar matrix?
```{r dat_2-twostep-1-RVE}
# meta regression with SE
rma.mv(g ~ 1 + sei,
    V = var.g,
    random = list( ~ 1 | Experiment, ~ 1 | ID),
    data = dat_2,
    method = "REML")
```

Here there is no evidence of small study effects so we don't need to model sampling variance.

Here I include the final model under the robust variance estimator.
```{r dat_2-final-RVE}
Q2m4.R  <-
  rma.mv(
    g ~ Trait.type:Host.type.2:Gradient.category -1,
    V = var.g,
    random = list( ~ 1 | ID, ~ 1 | Experiment),
    data = dat_2,
    method = "REML"
  )

# here we need to remove Resource vertebrate prevalence as there are not enough studies to estimate this effect robustly
Q2m4.R_est <- clubSandwich::coef_test(Q2m4.R, vcov = "CR2")[-12,]

forest.default(x= Q2m4.R_est$beta, sei =  Q2m4.R_est$SE, ci.lb = (Q2m4.R_est$beta - 1.96 * Q2m4.R_est$SE) , ci.ub =(Q2m4.R_est$beta + 1.96 * Q2m4.R_est$SE), 
                annotate=TRUE, showweights=T, header=F, 
                slab = c("Environment invertebrate intensity", "Environment invertebrate prevalence", "Environment vertebrate intensity", "Environment vertebrate prevalence", "Pollution invertebrate intensity", "Pollution invertebrate prevalence", "Pollution vertebrate intensity", "Pollution vertebrate prevalence", "Resource invertebrate intensity", "Resource invertebrate prevalence", "Resource vertebrate intensity"))
```

#### Data set 3: resistance and fitness effects of environmental stress
For now, I'm using the best model for Question 3: Q3m7
```{r dat_3-twostep-1}
# get standard errors of effect sizes 
dat_3$sei <- sqrt(dat_3$var.g)

# meta regression with SE
rma.mv(g ~ 1 + sei,
    V = varcovmat_3_PD$mat,
    random = list( ~ 1 | Experiment, ~ 1 | ID),
    data = dat_3,
    method = "REML")
```

Without the var-covar matrix
```{r dat_3-twostep-1-RVE}

# meta regression with SE
rma.mv(g ~ 1 + sei,
    V = var.g,
    random = list( ~ 1 | Experiment, ~ 1 | ID),
    data = dat_3,
    method = "REML")
```

As in data set 1, effects of small studies are dominated by negative effect sizes. Because the slope of sei is significant the variance is a better moderator for the overall effect.

Including the variance-covariance error matrix within studies
```{r dat_3-PB-adjusted}
Q3m7.B <-
  rma.mv(
    g ~ Trait.type:Gradient.category:Host.type.2 + var.g -1,
    V = varcovmat_3_PD$mat,
    random = list( ~ 1 | ID, ~ 1 | Experiment),
    data = dat_3,
    method = "REML"
  )

summary(Q3m7.B)

forest.default(x= Q3m7.B$beta, sei =  Q3m7.B$se, ci.lb = Q3m7.B$ci.lb, ci.ub = Q3m7.B$ci.ub, 
                annotate=TRUE, showweights=T, header=F, 
                slab = c("variance", "Environment invertebrate fecundity", "Environment invertebrate intensity", "Environment invertebrate prevalence", "Environment invertebrate survivorship", "Environment vertebrate intensity", "Environment vertebrate prevalence", "Environment vertebrate survivorship", "Pollution invertebrate fecundity", "Pollution invertebrate intensity", "Pollution invertebrate prevalence", "Pollution invertebrate survivorship", "Pollution vertebrate intensity", "Pollution vertebrate prevalence", "Pollution vertebrate survivorship", "Resource invertebrate fecundity", "Resource invertebrate intensity", "Resource invertebrate prevalence", "Resource invertebrate survivorship", "Resource vertebrate intensity", "Resource vertebrate prevalence", "Resource vertebrate survivorship"))
```

Using the robust variance estimator instead of the variance-covariance error matrix within studies
```{r dat_3-PB-adjusted-RVE}
Q3m7.B.R <-
  rma.mv(
    g ~ Trait.type:Gradient.category:Host.type.2 + var.g -1,
    V = var.g,
    random = list( ~ 1 | ID, ~ 1 | Experiment),
    data = dat_3,
    method = "REML"
  )

Q3m7.B.R_est <- clubSandwich::coef_test(Q3m7.B.R, vcov = "CR2")

forest.default(x= Q3m7.B.R_est$beta, sei =  Q3m7.B.R_est$SE, ci.lb = (Q3m7.B.R_est$beta - 1.96 * Q3m7.B.R_est$SE) , ci.ub =(Q3m7.B.R_est$beta + 1.96 * Q3m7.B.R_est$SE), 
                annotate=TRUE, showweights=T, header=F, 
                slab = c("variance","Environment invertebrate fecundity", "Environment invertebrate intensity", "Environment invertebrate prevalence", "Environment invertebrate survivorship", "Environment vertebrate intensity", "Environment vertebrate prevalence", "Environment vertebrate survivorship", "Pollution invertebrate fecundity", "Pollution invertebrate intensity", "Pollution invertebrate prevalence", "Pollution invertebrate survivorship", "Pollution vertebrate intensity", "Pollution vertebrate prevalence", "Pollution vertebrate survivorship", "Resource invertebrate fecundity", "Resource invertebrate intensity", "Resource invertebrate prevalence", "Resource invertebrate survivorship", "Resource vertebrate intensity", "Resource vertebrate prevalence", "Resource vertebrate survivorship"))
```

### Weighting the effects of (small) studies
A related question that one could want to check is how different studies contribute to the overall effects to see. In other words is our main result being driven by a few studies? Here I compute and plot a few measures of outlier/influence diagnostics.

I'm also calculating the influence statistics cooks distance and dfbetas at the level of studies. In the metafor man pages: 

"Cook's distance can be interpreted as the Mahalanobis distance between the entire set of predicted values once with the th case included and once with the th case excluded from the model fitting." 

"the DFBETAS value(s) essentially indicate(s) how many standard deviations the estimated coefficient(s) change(s) after excluding the th case from the model fitting."

A value is influential if "The hat value is larger than 3(/)." where p is the number of coefficients and k the number of cases.

#### Data set 1: fitness effects of environmental stress
I'm running the analysis here for the model with the lowest AIC (Q1m7). We should probably do it for whichever model is our main model in the manuscript (probably this one?)

```{r dat_1-influence}
# Cook's distance
Cook_dat1 <- cooks.distance(Q1m7, progbar=TRUE, dat_1$Study, reestimate=TRUE, parallel="multicore", ncpus = 2)

# how are these distances distributes?
hist(Cook_dat1)

# Which are the most influential studies?
Cook_dat1[which(Cook_dat1 > 1)]

# dfbetas
dfbetas_dat1 <- dfbetas.rma.mv(Q1m7, progbar=TRUE, dat_1$Study, reestimate=TRUE, parallel="multicore")

# Which are the most influential studies?
dfbetas_dat1[abs(dfbetas_dat1[,1])>1 | abs(dfbetas_dat1[,2])>1
             | abs(dfbetas_dat1[,3])>1 | abs(dfbetas_dat1[,4])>1
             | abs(dfbetas_dat1[,5])>1 | abs(dfbetas_dat1[,6])>1,]

# hat values
hat_dat1 <- hatvalues(Q1m7) 

# Which are the most influential effects?
inf_cutoff <- 3 * length(Q1m7$beta)/nrow(dat_1)

dat_1$Study[hat_dat1 > inf_cutoff]
```

#### Data set 2: resistance effects of environmental stress
I'm running the analysis here for the model with the lowest AIC (Q2m4). We should probably do it for which ever model is our main model in the manuscript (probably this one?)

```{r dat_2-influence}
# Cook's distance
Cook_dat2 <- cooks.distance(Q2m4, progbar=TRUE, dat_2$Study, reestimate=TRUE, parallel="multicore", ncpus = 2)

# how are these distances distributes?
hist(Cook_dat2)

# Which are the most influential studies?
Cook_dat2[which(Cook_dat2 > 1)]

# dfbetas
dfbetas_dat2 <- dfbetas.rma.mv(Q2m4, progbar=TRUE, dat_2$Study, reestimate=TRUE, parallel="multicore")

# Which are the most influential studies?
dfbetas_dat2[abs(dfbetas_dat2[,1])>1 | abs(dfbetas_dat2[,2])>1
             | abs(dfbetas_dat2[,3])>1 | abs(dfbetas_dat2[,4])>1
             | abs(dfbetas_dat2[,5])>1 | abs(dfbetas_dat2[,6])>1,]

# hat values
hat_dat2 <- hatvalues(Q2m4) 

# Which are the most influential effects?
inf_cutoff <- 3 * length(Q2m4$beta)/nrow(dat_2)

dat_2$Study[hat_dat2 > inf_cutoff]
```

#### Data set 3: resistance vs fitness effects of environmental stress
I'm running the analysis here for the model with the lowest AIC (Q3m4). We should probably do it for which ever model is our main model in the manuscript (probably this one?)

```{r dat_3-influence}
# Cook's distance
Cook_dat3 <- cooks.distance(Q3m4, progbar=TRUE, dat_3$Study, reestimate=TRUE, parallel="multicore", ncpus = 4)

# how are these distances distributes?
hist(Cook_dat3)

# Which are the most influential studies?
Cook_dat3[which(Cook_dat3 > 3)]

# dfbetas
dfbetas_dat3 <- dfbetas.rma.mv(Q3m4, progbar=TRUE, dat_3$Study, reestimate=TRUE, parallel="multicore")

# Which are the most influential studies?
dfbetas_dat3[abs(dfbetas_dat3[,1])>1 | abs(dfbetas_dat3[,2])>1
             | abs(dfbetas_dat3[,3])>1 | abs(dfbetas_dat3[,4])>1
             | abs(dfbetas_dat3[,5])>1 | abs(dfbetas_dat3[,6])>1,]

# hat values
hat_dat3 <- hatvalues(Q3m4) 

# Which are the most influential effects?
inf_cutoff <- 3 * length(Q3m4$beta)/nrow(dat_3)

dat_3$Study[hat_dat3 > inf_cutoff]
```

# Heterogeneity

Here I take the "best" model for each question to estimate between study heterogeneity. We could potentially do the same a different model or more models, depending on our goals.

Heterogeneity is important to know if the pooled effect size can be interpreted in a meaningful way. High heterogeneity is not necessarily a bad thing, but we would at least want to know if individual study effects usually go in the direction of the pooled effect.

## I2 statistic
Is one of the most commonly reported measures of heterogeneity. It captures the "percentage of variability in the effect sizes that is not caused by sampling error". That is, heterogeneity in true effect sizes. It tells us how much true heterogeneity there is *relative* to sampling error.

### I2 formula

```{r}
#' @title Parametric simulation 
#' @description Function for calculating I2 estimates using parametric simulations of model estimates taken from metafor. Note that the effectiveness of these simulations depends on the accuracy of model variance estimates.
#' @param estimate The estimate (i.e. variance) from a metafor model
#' @param sims The number of simulations 
#' @param n The sample size used in estimating the variance 
#' @author Daniel Noble - daniel.noble@anu.edu.au
#' @export
simMonteCarlo <- function(estimate, n, sims){
  tmp <- data.frame(num = base::rep(1:sims, each = n), 
                    y = stats::rnorm(n*sims, 0, base::sqrt(estimate)))
  Var <- tmp %>% dplyr::group_by(num) %>% dplyr::summarise(Mean_var = stats::var(y))
  return(as.numeric(Var$Mean_var))
}

## NOTE about PIPE: Run usethis::use_pipe() in the console. The package usethis will add what you need to import the pipe to your NAMESPACE and it will also drop warnings in checks

# Function for  rounding a data frame
round_df <- function(x, digits) {
  numeric_columns <- sapply(x, class) == 'numeric'
  x[numeric_columns] <-  round(x[numeric_columns], digits)
  x
}

# Function for estimating I2 
I2 <- function(model, v, ME = FALSE, sims = 1500, phylo = FALSE, obs = FALSE){
  
  if(class(model) != "rma.mv" && class(model) != "rma"){
    stop("The model object is not of class 'metafor'")
  }
  
  wi <- 1/v  #weight
  Vw <- sum((wi) * (length(wi) - 1))  / (((sum(wi)^2) - sum((wi)^2)))
  
  if("rma.mv" %in% class(model) | "rma" %in% class(model)){
    # Monte Carlo Simulations
    # From metafor extract the important statistics
    sigma2 <- matrix(model$sigma2, nrow = 1, ncol = length(model$sigma2))
    colnames(sigma2) <- model$s.names
    sigmaN <- model$s.nlevels
    
    if(obs == FALSE){
      stop("Please add the name of the observation-level random effect, obs. If models do not include this, re-run models including (~1|obs) in the random effect list")
    }
    
    #For each variance estimate use Monte Carlo simulation of data
    Sims <- data.frame(mapply(function(x,y) simMonteCarlo(x, y, sims = sims), x = sigma2, y = sigmaN))
    colnames(Sims) <- colnames(sigma2) 
    
    #Calculate total variance
    VT <- rowSums(cbind(Sims, Vw))
    Vt <- rowSums(Sims)  # remove Vw
    
    # For each variance component divide by the total variance. Note this needs to be fixed for phylo, but does deal with variable random effects.
    I2_re <- Sims / VT
    I2_total <- data.frame(Vt / VT)
    
    tmpMatrix <- data.frame(I2_re[, -match("ID", colnames(I2_re))], total = I2_total)
    names(tmpMatrix) = c(colnames(I2_re)[!colnames(I2_re) %in% 'ID'], 'total')
      
    CI <- lapply(tmpMatrix, function(x) stats::quantile(x, c(0.025, 0.975), na.rm = TRUE))
    I_CI <- as.data.frame(do.call(rbind, CI))
    colnames(I_CI) <- c("2.5% CI", "97.5% CI")
    I2_table <- cbind(I2_Est. = colMeans(tmpMatrix), I_CI )
    
    class(I2_table) <- c("metaAidR", "data.frame")
    
    return(round_df(I2_table, digits = 4))
  }
}
```


### Question 1
First  I calculate I2 using the formula above from Nakagawa and Santos (2012) for the model without the var-covar matrix

```{r  dat_1-I2-RVE}
I2.Q1m7 <- I2(Q1m7.B.R, v = dat_1$var.g, obs = "ID")*100
I2.Q1m7
```

If the variance-covariance matrix is included following the metafor example, total I2 becomes nearly 100%
```{r dat_1-I2}
# we account for covariance in sampling errors by using V, but assuming equal heterogeneity between fecundity vs survival and between environment, pollution and resources.

# matrix of sampling variances and covariances W 
W <- solve(varcovmat_1_PD$mat)

# model matrix
X <- model.matrix(Q1m7.B)

# estimate heterogeneity
P <- W - W %*% X %*% solve(t(X) %*% W %*% X) %*% t(X) %*% W

# overall I2: how much of total variance is attributed to heterogeneity
100 * sum(Q1m7.B$sigma2) / (sum(Q1m7.B$sigma2) + (Q1m7.B$k-Q1m7.B$p)/sum(diag(P)))

# Heterogeneity within experiments, between experiments, between parasites
100 * Q1m7.B$sigma2 / (sum(Q1m7.B$sigma2) + (Q1m7.B$k-Q1m7$p)/sum(diag(P)))

sav <- confint(Q1m7.B)

# confidence intervals for I2
100 * sav[[1]]$random[1,2:3] /(sum(Q1m7.B$sigma2) + (Q1m7.B$k-Q1m7.B$p)/sum(diag(P))) ### CI for ID-level I^2
100 * sav[[2]]$random[1,2:3] / (sum(Q1m7.B$sigma2)+ (Q1m7.B$k-Q1m7.B$p)/sum(diag(P))) ### CI for the experiment-level I^2
#100 * sav[[3]]$random[1,2:3] / (sum(Q1m7$sigma2) + (Q1m7$k-Q1m7$p)/sum(diag(P)))  ### CI for the parasite-level I^2
```

### Question 2
First  I calculate I2 using the formula above from Nakagawa and Santos (2012) for the model without the var-covar matrix

```{r  dat_2-I2-RVE}
I2.Q2m4 <- I2(Q2m4.R, v = dat_2$var.g, obs = "ID")*100
I2.Q2m4
```

If the variance-covariance matrix is included following the metafor example, total I2 becomes nearly 100%
```{r dat_2-I2}
# we account for covariance in sampling errors by using V, but assuming equal heterogeneity between fecundity vs survival and between environment, pollution and resources.

# matrix of sampling variances and covariances W 
W <- solve(varcovmat_2_PD$mat)

# model matrix
X <- model.matrix(Q2m4)

# estimate heterogeneity
P <- W - W %*% X %*% solve(t(X) %*% W %*% X) %*% t(X) %*% W

# overall I2: how much of total variance is attributed to heterogeneity
100 * sum(Q2m4$sigma2) / (sum(Q2m4$sigma2) + (Q2m4$k-Q2m4$p)/sum(diag(P)))

# Heterogeneity between experiments, within experiments, between parasites
100 * Q2m4$sigma2 / (sum(Q2m4$sigma2) + (Q2m4$k-Q2m4$p)/sum(diag(P)))

sav2 <- confint(Q2m4)

# confidence intervals for I2
100 * sav2[[1]]$random[1,2:3] /(sum(Q2m4$sigma2) + (Q2m4$k-Q2m4$p)/sum(diag(P))) ### CI for experiment-level I^2
100 * sav2[[2]]$random[1,2:3] / (sum(Q2m4$sigma2)+ (Q2m4$k-Q2m4$p)/sum(diag(P))) ### CI for the ID-level I^2
#100 * sav2[[3]]$random[1,2:3] / (sum(Q2m4$sigma2) + (Q2m4$k-Q2m3$p)/sum(diag(P)))  ### CI for the parasite-level I^2
```

### Question 3
First  I calculate I2 using the formula above from Nakagawa and Santos (2012) for the model without the var-covar matrix

```{r  dat_3-I2-RVE}
I2.Q3m7 <- I2(Q3m7.B.R, v = dat_3$var.g, obs = "ID")*100
I2.Q3m7
```

If the variance-covariance matrix is included following the metafor example, total I2 becomes nearly 100%

```{r dat_3-I2}
# we account for covariance in sampling errors by using V, but assuming equal heterogeneity between fecundity vs survival and between environment, pollution and resources.

# matrix of sampling variances and covariances W 
W <- solve(varcovmat_3_PD$mat)

# model matrix
X <- model.matrix(Q3m7.B)

# estimate heterogeneity
P <- W - W %*% X %*% solve(t(X) %*% W %*% X) %*% t(X) %*% W

# overall I2: how much of total variance is attributed to heterogeneity
100 * sum(Q3m7.B$sigma2) / (sum(Q3m7.B$sigma2) + (Q3m7.B$k-Q3m7.B$p)/sum(diag(P)))

# Heterogeneity between experiments, within experiments, between parasites
100 * Q3m7.B$sigma2 / (sum(Q3m7.B$sigma2) + (Q3m7.B$k-Q3m7.B$p)/sum(diag(P)))

sav3 <- confint(Q3m7.B)

# confidence intervals for I2
100 * sav3[[1]]$random[1,2:3] /(sum(Q3m7.B$sigma2) + (Q3m7.B$k-Q3m7.B$p)/sum(diag(P))) ### CI for ID-level I^2
100 * sav3[[2]]$random[1,2:3] / (sum(Q3m7.B$sigma2)+ (Q3m7.B$k-Q3m7.B$p)/sum(diag(P))) ### CI for the experiment-level I^2
#100 * sav3[[3]]$random[1,2:3] / (sum(Q3m7.B$sigma2) + (Q3m7.B$k-Q3m7.B$p)/sum(diag(P)))  ### CI for the parasite-level I^2
```

## Visualizations
Orchard plots are a useful visualization technique for the distribution of effect size  (Nakagawa et al 2020). These plots cannot accommodate var-covar error matrices or covariates. For now, I will exclude host type and look at effects of each gradient category and trait type.

### Question 1

```{r Q1-orchard, message=FALSE}
Q1m7.1 <-
  rma.mv(
    g ~ Gradient.category:Trait.type -1,
    V = var.g,
    random = list( ~ 1 | ID, ~ 1 | Experiment),
    data = dat_1,
    method = "REML"
  )

pal <- RColorBrewer::brewer.pal(6,"Spectral")
orchard_plot(Q1m7.1, mod = "var.g", xlab = "Standardised mean difference", transfm = "none", angle = 45, alpha = 0.7)+
 scale_fill_manual(values = pal) + scale_colour_manual(values = pal) #+ 
scale_y_discrete( labels= c("Environment:Fecundity", "Pollution:Fecundity",  "Resource:Fecundity","Environment:Survival", "Pollution:Survival", "Resource:Survival"))
```

### Question 2

```{r Q2-orchard, message=FALSE}
Q2m4.1 <-
  rma.mv(
    g ~ Gradient.category:Trait.type -1,
    V = var.g,
    random = list( ~ 1 | ID, ~ 1 | Experiment),
    data = dat_2,
    method = "REML"
  )

pal <- RColorBrewer::brewer.pal(6,"Spectral")
orchard_plot(Q2m4.1, mod = "Gradient.category", xlab = "Standardised mean difference", transfm = "none", angle = 45, alpha = 0.7)+
 scale_fill_manual(values = pal) + scale_colour_manual(values = pal) + 
scale_y_discrete( labels= c("Environment:Intensity", "Pollution:Intensity",  "Resource:Intensity","Environment:Prevalence", "Pollution:Prevalence", "Resource:Prevalence")) +
  xlim(-5,15.5)
```

### Question 3
We may want to consider changing the order in the y axis.
```{r Q3-orchard, message=FALSE}
Q3m7.1 <-
  rma.mv(
    g ~ Gradient.category:Trait.type -1,
    V = var.g,
    random = list( ~ 1 | ID, ~ 1 | Experiment),
    data = dat_3,
    method = "REML"
  )

pal <- colorRampPalette(brewer.pal(11,"Spectral"))(length(Q3m7.1$beta))
orchard_plot(Q3m7.1, mod = "Gradient.category", xlab = "Standardised mean difference", transfm = "none", angle = 45, alpha = 0.7)+
 scale_fill_manual(values = pal) + scale_colour_manual(values = pal) + 
scale_y_discrete( labels= c("Environment:Fecundity", "Pollution:Fecundity",  "Resource:Fecundity","Environment:Intensity", "Pollution:Intensity",  "Resource:Intensity","Environment:Prevalence", "Pollution:Prevalence", "Resource:Prevalence","Environment:Survival", "Pollution:Survival", "Resource:Survival")) 
```
