---
title: "Host-Parasite Meta-analysis"
author: "Dave Civitello, Beatriz Willink, Amanda Vicente"
date: "3/3/2021"
output: html_document
---

Read in data and load packages
```{r, message=FALSE}
dat <- read.csv("../data/MA_test_dat2.csv", header = T, sep = ",")
x <-
  c(
    "metafor")

lapply(x, function(y) {
  # check if installed, if not install
  if (!y %in% installed.packages()[, "Package"])
    install.packages(y)
  
  # load package
  try(require(y, character.only = T), silent = T)
})
```


# Getting effect sizes

## Calculate OR and approximate variance from contingency tables
```{r}
# store sample size in an object
N <- nrow(dat)

for (i in 1:N) {
  # variance cannot be estimated if all individuals in a category are 0
  if (dat$Variation.Type[i] == "OR" &
      dat$Success_C[i] > 0 & dat$Success_X[i] > 0 & (dat$N_C[i] - dat$Success_C[i]) > 0 & (dat$N_X[i] - dat$Success_X[i]) > 0) {
    dat$OR[i] <-
      (dat$Success_X[i] * (dat$N_C[i] - dat$Success_C[i])) / (dat$Success_C[i] * (dat$N_X[i] - dat$Success_X[i]))
    dat$Log.OR[i] <- log(dat$OR[i])
    # approximate variance
    dat$Log.OR.v[i] <-
      1 / dat$Success_X[i] + 1 / (dat$N_X[i] - dat$Success_X[i]) + 1 / dat$Success_C[i] + 1 / (dat$N_C[i] - dat$Success_C[i])
  }
  else {
    # if there is no appropiate data, make these columns NA
    dat$OR[i] <- NA
    dat$Log.OR[i] <- NA
    dat$Log.OR.v[i] <- NA
  }
}

```


## Calculate SD for comparisons with continuous normally distributed variables
** What do we do with CIs from non-normal distributions?**
```{r}
# create SD variables
dat$SD_C <- vector(length = N)
dat$SD_X <- vector(length = N)


for (i in 1:N) {
  
  # Calculate SD if SE is reported
  if (dat$Variation.Type[i] == "SE") {
    dat$SD_C[i] <- dat$Variation_C[i] * sqrt(dat$N_C[i])
    dat$SD_X[i] <- dat$Variation_X[i] * sqrt(dat$N_X[i])
  } else {
   
     # calculate SD if a 95% confidence interval for a normal distribution is reported
    # this also applies to the Wald confidence interval for prorportions, as it is a normal approximation to the binomial
    if (dat$Variation.Type[i] == "95CI" |
        dat$Variation.Type[i] == "Confidence limit (Wald CI)") {
     
       # calculate SE from lower and upper limits
      temp_C1 <- (dat$Upr_C[i] - dat$Mean_C[i]) / 1.96
      temp_C2 <- (dat$Mean_C[i] - dat$Lwr_C[i]) / 1.96
      
      temp_X1 <- (dat$Upr_X[i] - dat$Mean_X[i]) / 1.96
      temp_X2 <- (dat$Mean_X[i] - dat$Lwr_X[i]) / 1.96
      
      # average digitized/recorded values of SE (because we have two) and transform to SD
      dat$SD_C[i] <- mean(abs(c(temp_C1, temp_C2))) * sqrt(dat$N_C[i])
      dat$SD_X[i] <- mean(abs(c(temp_C1, temp_C2))) * sqrt(dat$N_X[i])
    } else {
     
       # if SD is reported leave as such
      if (dat$Variation.Type[i] == "SD") {
        dat$SD_C[i] <- dat$Variation_C[i]
        dat$SD_X[i] <- dat$Variation_X[i]
      } else {
        
        # if there is no appropiate data, make these columns NA
        dat$SD_C[i] <- NA
        dat$SD_X[i] <- NA
      }
    }
  }
}
```


## Get effect sizes and variances
```{r}
# within groups standard deviation
dat$S_within <-
  sqrt(((dat$N_C - 1) * dat$SD_C ^ 2 + (dat$N_X - 1) * dat$SD_X ^ 2) / (dat$N_C + dat$N_X - 2))

# standardized effect size
for (i in 1:N) {
  # standardized mean difference
  # we can't include data without variance
  if (is.na(dat$S_within[i]) == FALSE & dat$S_within[i] > 0) {
    dat$d[i] <- (dat$Mean_X[i] - dat$Mean_C[i]) / dat$S_within[i]
    dat$var.d[i] <-
      (dat$N_C[i] + dat$N_X[i]) / (dat$N_C[i] * dat$N_X[i]) + (dat$d[i] ^ 2) /
      (2 * (dat$N_C[i] + dat$N_X[i]))
    
  } else{
    # if only F statistic is reported
    if (dat$Variation.Type[i] == "F_X") {
      dat$d[i] <-
        sqrt(dat$F_X[i] * (dat$N_X[i] + dat$N_C[i]) / (dat$N_C[i] * dat$N_X[i]))
      dat$var.d[i] <-
        (dat$N_C[i] + dat$N_X[i]) / (dat$N_C[i] * dat$N_X[i]) + (dat$d[i] ^ 2) /
        (2 * (dat$N_C[i] + dat$N_X[i]))
      
    } else{
      # converting log OR
      if (is.na(dat$Log.OR.v[i]) == FALSE & dat$Log.OR.v[i] > 0) {
        dat$d[i] <- dat$Log.OR[i] * (sqrt(3) / pi)
        dat$var.d[i] <- dat$Log.OR.v[i] * 3 / pi ^ 2
      } else {
        # for now, leave other types of effects as NA
        dat$d[i] <- NA
        dat$var.d[i] <- NA
      }
    }
  }
}
```

## Correct for sample size and get g
```{r}
# sample size correction factor
dat$J <- 1 - 3 / (4 * (dat$N_C + dat$N_X - 2) -1)

# corrected effect size
dat$g <- dat$J * dat$d

# and variance
dat$var.g <- (dat$J ^ 2) * dat$var.d
```

# preliminary analysis
**!!!! Not final analysis, not for publication !!!!**
```{r}
m1 <-
  rma.mv(
    g ~ 1,
    V = var.g,
    random = list( ~ 1 |
                     ID, ~ 1 | Experiment, ~ 1 | Parasite),
    data = dat,
    method = "REML"
  )

summary(m1)

m2 <-
  rma.mv(
    g ~ Trait.category - 1,
    V = var.g,
    random = list( ~ 1 |
                     ID, ~ 1 | Experiment,  ~ 1 | Parasite),
    data = dat,
    method = "REML"
  )

summary(m2)
```

