---
title: "Disease across gradients"
author: "DJC"
date: "1/26/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Disease dynamics across environmental gradients, based on Amanda's meta-analysis

Gonna set up models to run over gradients of [Env factor] or an SI-Resource model and then embed the results that Amanda found in her meta-analysis of how traits respond to this variation.

### Effects that Amanda detected in the meta-analysis

First, tested for effects of stressors on demographic traits of uninfected vs. infected hosts (are infected more sensitive)? Meta-analysis found no *proportional* difference between the groups for survival or reproduction. So in the models, when we incorporate these effects, we want to a common parameter for the strength of these effects on both groups.

#### Low resources
Have negative effect on fecundity
Causes a decrease in intensity


#### Environmental gradients
Have negative effect on survivorship
Bad environments cause an increase in intensity

#### Toxicants
Have negative effect on survivorship
Bad environments cause a decrease in prevalence

```{r}
rtools <- "C:\\Rtools\\bin"
gcc <- "C:\\Rtools\\gcc-4.6.3\\bin"
path <- strsplit(Sys.getenv("PATH"), ";")[[1]]
new_path <- c(rtools, gcc, path)
new_path <- new_path[!duplicated(tolower(new_path))]
Sys.setenv(PATH = paste(new_path, collapse = ";"))

# compile food gradient model from C definition
try(dyn.unload("C:/RData/food_gradient.dll")) # unload dll
system("R CMD SHLIB C:/RData/food_gradient.c")
dyn.load("C:/RData/food_gradient.dll") # Load dll

# compile environmental gradient model from C definition
try(dyn.unload("C:/RData/env_gradient.dll")) # unload dll
system("R CMD SHLIB C:/RData/env_gradient.c")
dyn.load("C:/RData/env_gradient.dll") # Load dll

```



```{r models}
library(deSolve)
library(ggplot2)

Inits = c(S = 10, I = 1, R = 10)
pars = c(r = 1, K = 10, fM = 0.1, hr = 4, ht = 2, e = 0.5, rho = 0.25, BM = 0.01, d = 0.01, v = 0.04)
timespan = 100000

sim1 = lsoda(y = Inits, times=0:timespan, parms = pars, func="derivs", dllname = "food_gradient", initfunc="initmod",  maxsteps=1e6)
#plot(I ~ time, data=sim1)

# Simulate over resource growth rate gradient and collect results
S_result = numeric()
I_result = numeric()
R_result = numeric()
r_inputs = seq(from = 0.1, to = 1, length.out = 100)
ht_inputs = c(0, 1, 2, 4, 8)

r_sim_inputs = numeric()
ht_sim_inputs = numeric()
k = 1
    
for(i in 1:length(r_inputs)){
  pars["r"] = r_inputs[i]
  for(j in 1:length(ht_inputs)){
    pars["ht"] = ht_inputs[j]
    sim = lsoda(y = Inits, times=0:timespan, parms = pars, func="derivs", dllname = "food_gradient", initfunc="initmod", hmax=0.1,  maxsteps=1e7)
    r_sim_inputs[k] = r_inputs[i]
    ht_sim_inputs[k] = ht_inputs[j]
    S_result[k] = sim[dim(sim)[1], 2]
    I_result[k] = sim[dim(sim)[1], 3]
    R_result[k] = sim[dim(sim)[1], 4]
    if(k %% 1000 == 0){print(k)}
    k = k + 1
  }
}

sim_summary = data.frame("resource_r" = r_sim_inputs, "ht" = ht_sim_inputs, "S_eq" = S_result, "I_eq" = I_result, "R_eq" = R_result,
                         "Prev_eq" = I_result/(S_result + I_result))

plot1 <- ggplot(sim_summary, aes(x = resource_r, y = Prev_eq, group = ht)) +
          geom_line() 
plot1

```
```{r models2}
library(deSolve)
library(ggplot2)

SI_Env = function(t, y, parameters) {
  S=y[1]; I=y[2];                 # For convenience, I like to spell out which state varible within the vector y is which
  with(as.list(parameters),{         # The with() function allows us to write very clean equations for the model by telling r where to look for parameter names
    dSdt= b*(S + rho*I)*(1 - (S+I)/K) - B*exp(BE*ENV)*S*I - d*exp(dE*ENV)*S    
    dIdt= B*exp(BE*ENV)*S*I - (d + v)*exp(dE*ENV)*I   
    result = c(dSdt,dIdt)          # The function for simulation expects us to store the result this way
    return(list(result))
  }
  ) 
}

Inits = c(S = 10, I = 1)
pars = c(b = 0.5, rho = 0.25, K = 10, B = 0.01, BE=1, d = 0.01, dE=2, ENV=0, v = 0.04)
timespan = 10000

sim1 = ode(y=Inits, times=0:timespan, method="lsoda", func=SI_Env, parms = pars)

# Simulate over resource growth rate gradient and collect results
S_result = numeric()
I_result = numeric()
E_inputs = seq(from = 0, to = 1, length.out = 100)
BE_inputs = c(-8, -4, -2, 0, 2, 4, 8)#seq(from = -1, to = 40, length.out = 100)

Env_inputs = numeric()
BE_sim_inputs = numeric()
k = 1
    
for(i in 1:length(E_inputs)){
  pars["ENV"] = E_inputs[i]
  for(j in 1:length(BE_inputs)){
    pars["BE"] = BE_inputs[j]
    sim = lsoda(y = Inits, times=0:timespan, parms = pars, nout=1, func="derivs", dllname = "env_gradient", initfunc="initmod",  maxsteps=1e6)
    Env_inputs[k] = E_inputs[i]
    BE_sim_inputs[k] = BE_inputs[j]
    S_result[k] = sim[dim(sim)[1], 2]
    I_result[k] = sim[dim(sim)[1], 3]
    if(k %% 1000 == 0){print(k)}
    k = k + 1
  }
}

sim_summary = data.frame("Environment" = Env_inputs, "Beta_sensitivity" = BE_sim_inputs, "S_eq" = S_result, "I_eq" = I_result, "Prev_eq" = I_result/(S_result + I_result))

plot1 <- ggplot(sim_summary, aes(x = Environment, y = Prev_eq, group = Beta_sensitivity)) +
          geom_line() 
plot1


```